<!-- Mandatory section, title may vary. Core of the report. -->

# Analysis and response

In this Science Response, all figure pages were rebuilt with updated data using the methods described in @anderson2019synopsis.
Survey data from 2021, 2020, 2019, and in one case, 2018, were added (Table 1).
Commercial data from 2021, 2020, and 2019 were added.

```{r survey-years-table, results='asis'}
# x <- readRDS("../gfsynopsis/report/data-cache-old?/longnose-skate.rds")
# x <- x$survey_index
# group_by(x, survey_abbrev) |>
#   select(survey_abbrev, year) |>
#   readr::write_csv("report/2019-resdoc-survey-years.csv")

rd <- readr::read_csv(here::here("report/2019-resdoc-survey-years.csv"),
  show_col_types = FALSE
) |>
  mutate(rd_year = year)

sr <- readRDS(here::here("report/data-cache/longnose-skate.rds"))$survey_index
sr <- sr |> select(survey_abbrev, year)

surveys <- c(
  "SYN WCHG", "SYN HS", "SYN QCS", "SYN WCVI",
  "HBLL OUT N",
  "HBLL OUT S", "HBLL INS N", "HBLL INS S",
  "OTHER HS MSA", "IPHC FISS"
)
srvs <- tibble(survey_abbrev = surveys, order = seq_along(surveys))

yrs <- left_join(sr, rd) |>
  right_join(srvs) |>
  arrange(order) |>
  select(-order)
yrs <- filter(yrs, is.na(rd_year)) |>
  select(-rd_year)
yrs <- filter(yrs, survey_abbrev != "IPHC FISS")
yrs <- bind_rows(
  yrs,
  tibble(survey_abbrev = c("IPHC FISS", "IPHC FISS", "IPHC FISS"), year = c(2019, 2020, 2021))
)

csasdown::csas_table(yrs,
   col_names = c("Survey", "Years added"),
   caption = "Fisheries independent survey years added in this Science Response compared to the previous Anderson et al. (2019) report."
)
```


```{r child="02-iphc-update.Rmd", echo=FALSE, cache=FALSE}
```

## Caveats {#sec:caveats}

There are several caveats when interpreting this report as outlined in @anderson2019synopsis.
Primary caveats include:

1. The outputs in this report are not a substitute for stock assessment. For
   example, although relative biomass index trends from surveys indicate the
   biomass trend for a species in an area, such information is best combined
   with other information such as removals by commercial catches and
   information on the age- or length-composition of the stock to make
   conclusions about the status of a stock.

1. Biomass indices from trawl or longline surveys and commercial CPUE indices
   need careful interpretation on a stock-by-stock basis. We have attempted to
   flag survey index trends that may be especially suspect either because of
   high survey variability or because only a small fraction of trawl or
   longline sets contain the species, but this is not a guarantee in itself.
   Survey indices are not always representative of abundance for a variety of
   reasons, and a lack of data for a species does not necessarily indicate
   a small population---the surveys may simply not be suitable for sampling
   that species. Furthermore, changes through time, including fish behavioural
   changes or range shifts, could result in biases through time even for
   well-sampled species.

1. Survey and commercial CPUE index trends do not resolve population scale and
   the outputs in this report do not resolve conflicts in trends drawn from
   different sources for the same species.

1. The outputs in this report are not appropriate for marine spatial planning.
   The data as presented are resolved at a coarse spatial scale and marine
   spatial planning uses require specific data treatments beyond the general
   approaches used in this report.

1. The commercial CPUE data should not be considered to be proportional to
   stock abundance for a multitude of reasons [e.g., @harley2001]. Nonetheless,
   we think there is value in transparently displaying the available data for
   all species.

1. The catch history reported here reflects recorded data and may not represent
   actual catches. The commercial catch presented here will not necessarily
   match reconstructed time series in stock assessments. Historical catch
   reconstructions require careful species-specific consideration and analysis.
   Furthermore, fluctuations in commercial catch do not necessarily reflect
   declines in stock abundance and may be due to other factors including
   implementation of management measures. Reported discard weights are
   considered less reliable prior to 100% observer coverage of the bottom trawl
   fishery in 1996, and prior to fisheries integration in 2006 for the trap,
   hook and line, midwater trawl and Strait of Georgia bottom trawl fisheries.
   The discards in the catch plots therefore only include bottom trawl discard
   weights from 1996 to present and trap, hook and line, midwater trawl and
   Strait of Georgia bottom trawl discard weights from 2006 to present.

1. It is not feasible to individually assess the results for all species in
   a detailed manner. To use the results for a particular species in future
   assessments, or to make other inference, we recommend that users carefully
   examine the data and model results. Due to the necessary automation required
   to construct this report, not all species-specific special cases may have
   been fully considered.

In addition to these caveats noted in @anderson2019synopsis, coronavirus
disease 2019 (COVID-19) resulted in several cancelled surveys in 2020, some of
which were then conducted in 2021 in addition to the scheduled surveys.
Furthermore, COVID-19 caused several changes within the commercial fleet in
2020 and 2021. This affects the catch figures (e.g., Figure \@ref(fig:catches)) as well
as potentially affect the standardization process of the commercial catch per
unit effort (e.g., Figure \@ref(fig:trawl-cpue-index), Appendix D in Anderson et al. 2019).

## Data accessibility

Data in this document are maintained by the Groundfish Data Unit at the Pacific
Biological Station in Nanaimo, British Columbia. Data from the Synoptic Bottom
Trawl Surveys and the Hard Bottom Longline (HBLL) surveys are available through
the Canadian [Open Government Portal](https://open.canada.ca/data/en/dataset?portal_type=dataset)
by searching for 'groundfish synoptic' or 'HBLL'.
For the IPHC survey, [gfiphc](https://github.com/pbs-assess/gfiphc) contains
some data and has details for requesting the rest.
Requests for data held by the
DFO Pacific Region that are not available through the Open Government Portal
can be made as described on the
[Pacific Fisheries Catch Statistics](http://www.pac.dfo-mpo.gc.ca/stats/index-eng.html)
website.

## Reproducibility

All of the data extraction, data manipulation, model fitting, and visualization for this report is automated and reproducible.
The gfdata, gfplot, gfsynopsis, sdmTMB, and csasdown R packages were developed for this purpose.
The gfdata package enables the data extraction.
The gfplot package performs the model fitting and visualizations.
The gfsynopsis package calls functions from the gfplot and gfdata packages to generate this report.
The sdmTMB package fits spatially explicit generalized linear mixed effect models with random fields to produce the survey maps.
The csasdown package builds this report using R Markdown and bookdown.
The source code for this report is available on the GitHub repository [pbs-assess/gfsynopsis](https://github.com/pbs-assess/gfsynopsis/), although the report can only be built with access to the Pacific Biological Station network.

```{r sha}
sha <- list()
sha[[1]] <- system("cd ../../../gfdata; git rev-parse HEAD", intern = TRUE)
sha[[2]] <- system("cd ../../../gfplot; git rev-parse HEAD", intern = TRUE)
sha[[3]] <- system("cd ../../../csasdown; git rev-parse HEAD", intern = TRUE)
sha[[4]] <- system("cd ../../../sdmTMB; git rev-parse HEAD", intern = TRUE)
sha[[5]] <- system("cd ../../../gfiphc; git rev-parse HEAD", intern = TRUE)
sha[[6]] <- system("git rev-parse HEAD", intern = TRUE)
names(sha) <- c("gfdata", "gfplot", "csasdown", "sdmTMB", "gfiphc", "gfsynopsis")
sha <- lapply(sha, function(x) substr(x, 1, 6))
```

The specific versions used to generate this report are available at:

* [pbs-assess/gfdata](`r paste0("https://github.com/pbs-assess/gfdata/tree/", sha$gfdata)`)\
* [pbs-assess/gfplot](`r paste0("https://github.com/pbs-assess/gfplot/tree/", sha$gfplot)`)\
* [pbs-assess/gfiphc](`r paste0("https://github.com/pbs-assess/gfiphc/tree/", sha$gfiphc)`)\
* [pbs-assess/gfsynopsis](`r paste0("https://github.com/pbs-assess/gfsynopsis/tree/", sha$gfsynopsis)`)\
* [pbs-assess/sdmTMB](`r paste0("https://github.com/pbs-assess/sdmTMB/tree/", sha$sdmTMB)`)\
* [pbs-assess/csasdown](`r paste0("https://github.com/pbs-assess/csasdown/tree/", sha$csasdown)`)\

# Conclusions

The report presents an update to @anderson2019synopsis that adds data primarily from 2019 to 2021 as well as adding citations to any new stock assessment or Committee on the Status of Endangered Wildlife in Canada (COSEWIC) reports.
Temporal and spatial patterns in survey and commercial catch for species are evident by scanning the main figure pages.
This report is expected to be updated on approximately a biennial basis.
These updates will include any new data since the previous report and any important corrections to the data, text, or visualizations.
On a less frequent basis, the authors will consider making larger changes to the structure, methods, or content of the report within the context of a CSAS review process.
It is intended that this report will continue to (1) facilitate regular review by groundfish scientists and managers of trends in survey indices and stock composition across all species; (2) facilitate generating standardized datasets, visualizations, and biological model fits to help assessment scientists; and (3) increase data transparency between DFO, the fishing industry, First Nations, non-governmental organizations, and the general public.
