# PLOT DESCRIPTIONS {#sec:plot-descriptions}

```{r load-example-figs}
g <- readRDS(here::here("report/ggplot-objects/petrale-sole.rds"))
sp <- "Petrale Sole"
```

<!-- A fake square to use to get the aspect ratio of the maps correct: -->

```{r checking-square}
checking_square <- geom_polygon(data = data.frame(x = c(400, 600, 600, 400),
  y = c(5500, 5500, 5700, 5700)), aes_string(x = "x", y = "y"),
  inherit.aes = FALSE, fill = "grey50", lwd = 1, col = "black")
```

In this section we provide complete captions for each of the
visualizations that form the species-by-species figures in
Section \@ref(sec:synopsis-plots). We use `r sp` as an example species for all
plots except for commercial catch per unit effort maps where we use Pacific Cod.

## RELATIVE BIOMASS INDEX TRENDS FROM SURVEYS

```{r make-surv-abbrev-text, eval=TRUE}
# happens to have data for all the surveys:
dat_survey_index <- readRDS(here::here("report/data-cache/canary-rockfish.rds"))
dat_survey_index <- dat_survey_index$survey_index
survs <- c(
  "SYN WCHG", "SYN HS", "SYN QCS", "SYN WCVI",
  "HBLL OUT N",
  "HBLL OUT S", "HBLL INS N", "HBLL INS S",
  "MSA HS", "IPHC FISS")
survs <- tibble(surv_order = seq_along(survs), survey_abbrev = survs)
survey_descriptions <-
  semi_join(dat_survey_index, survs, by = "survey_abbrev") %>%
  select(survey_series_desc, survey_abbrev) %>%
  unique()
survey_descriptions <- bind_rows(survey_descriptions,
  tibble(survey_series_desc =
    "Hecate Strait Multispecies Assemblage Bottom Trawl",
     survey_abbrev = "MSA HS"))
x <- inner_join(survey_descriptions, survs, by = "survey_abbrev") %>%
  arrange(surv_order)
x$survey_series_desc <- gsub(" $", "", x$survey_series_desc)
surv_abbrev_text <- paste0(x$survey_abbrev, " = ", x$survey_series_desc, collapse = ", ")
```

(ref:survey-index-example) `r paste0("Example relative biomass index trends from trawl and longline surveys for ", sp, ". Dots represent mean estimates of relative biomass and shaded ribbons around the dots and lines represent 95\\% bootstrap confidence intervals. 'Mean CV' is the mean of the annual coefficients of variation (CVs), and 'Mean $+$ve sets' indicates the ratio of the mean number (across the years) of sets that captured the species of interest to the mean number of sets. Grey shaded panels indicate survey trends with 'Mean CV' greater than 0.4 or 'Mean $+$ve sets' less than 5\\%. All vertical axes are scaled between zero and the maximum upper confidence interval value for that survey. Time series with light grey dots and dotted upper and lower lines for 95% confidence intervals represent a further index that has been standardized with a geostatistical model. ", surv_abbrev_text, ". For the IPHC FISS, the values are relative counts per effective skate rather than biomass..")`

```{r survey-index, message=FALSE, warning=FALSE, fig.asp=0.9, fig.width=6, out.width="4.7in", fig.cap="(ref:survey-index-example)"}
g$survey_index
```

\clearpage

## MAPS OF RELATIVE BIOMASS FROM SURVEYS

```{r map-text}
map_text <- "Note that the coast has been rotated 40$^{\\circ}$ to fit all the maps in the available space. Depth contours are shown at 100 m, 200 m, and 500 m."
```

(ref:survey-maps-cap) Example maps of relative biomass (or catch rate) from
trawl and longline surveys from the latest available years of each survey for
`r sp`. Shown are the synoptic trawl surveys (left), the outside hard bottom long
line (HBLL OUT) surveys (middle), and the IPHC FISS (right). Individual sets
are shown in the two left panels as faint crosses (if the species was not
caught in that set), or circles with the area of the circle proportional to the
species density from the set. Colour shading indicates predictions from
a spatial model that includes depth and depth squared as a predictor as well
as spatial random effects (Appendix \@ref(app:spatial-modeling)). The colour scale
is fourth-root transformed to render a visual pattern similar to a log
transformation without overemphasizing differences close to zero. The colour
scale ('Viridis plasma') is perceptually uniform, robust to colour blindness,
and prints accurately in grayscale [@viridis2018]. The synoptic and HBLL maps
show predicted biomass density throughout the survey domain. The IPHC map shows
the raw unmodelled data for fixed station locations --- stations without any
observations for a given species are shown as empty circles. Years on the left
side of each plot indicate the year of the respective survey. Surveys (except
IPHC) in which less than 2\% of the sets contained the species are not modeled
and are shown with raw data only. Mean values shown at the bottom are the mean
fish density values from the raw data for the entire coast for the indicated years.
`r map_text`

```{r survey-maps, warning=FALSE, message=FALSE, fig.asp=1.13, fig.width=6, out.width="4.0in", fig.cap="(ref:survey-maps-cap)", fit.pos="H"}
trans <- "fourth_root_power"
units_transformation <- (1000 * 1000) / 1000
breaks <- c(0.002, 0.05, 0.3)
p1 <- g$survey_spatial_syn +
  labs(fill = "Predicted\nbiomass\ndensity\n(kg/km^2)", parse = TRUE) +
  ggplot2::scale_fill_viridis_c(trans = trans, option = "C", breaks = breaks, 
    labels = breaks * units_transformation) +
  guides(fill = guide_colorbar(), size = FALSE)

p2 <- g$survey_spatial_hbll +
  labs(fill = "Predicted\nbiomass\ndensity\n(fish/km^2)") +
  ggplot2::scale_fill_viridis_c(trans = trans, option = "C", breaks = c(2, 10, 50)) +
  guides(fill = guide_colorbar(), size = FALSE)

p3 <- g$survey_spatial_iphc +
  labs(colour = "Catch\nrate\n(fish/effective\nskate)") +
  guides(colour = guide_colorbar(), size = FALSE) +
  ggplot2::scale_fill_viridis_c(trans = "fourth_root_power", option = "C",
    na.value = 'white', breaks = c(0.2, 0.4, 0.7)) +
  ggplot2::scale_colour_viridis_c(trans = "fourth_root_power", option = "C",
    na.value = 'grey35', breaks = c(0.2, 0.4, 0.7)) +
  ggtitle("IPHC catch rate")

gridExtra::grid.arrange(
  p1,
  p2,
  p3, # + checking_square,
  nrow = 1)
```

\clearpage

## COMMERCIAL FISHERY CATCHES

<!-- % TODO: Might need to preface with the commercial groundfish area map or provide a link to groundfish webpage: http://www.pac.dfo-mpo.gc.ca/fm-gp/commercial/ground-fond/index-eng.html See also for North to South orientation of plots: coastwide, 5CDE, 5AB, 3CD. -->

(ref:catches-cap) Example commercial fishery catch plots for `r sp`.
Catch from various gear types is indicated by colour shading. Catch is
calculated as the summed weight of landings aggregated by year. Discards include
reported discard weights from all fisheries combined; however, bottom trawl
discards are considered less reliable prior to 100% observer coverage in 1996
and trap, hook and line, midwater trawl and Strait of Georgia bottom trawl
discards are less reliable prior to fisheries integration in 2006 and are
therefore not included. Years before 1996 and 2006 are shaded
grey to indicate that catches are considered less reliable than modern data. 
Foreign catches recorded in our databases between 1977 and 1991
Management areas, as indicated in the top left corner of each panel, are shown
in Figure \@ref(fig:management-map).

```{r catches, fig.asp=1.3, fig.width=5, out.width="4in", warning=FALSE, fig.cap="(ref:catches-cap)"}
g$catch
```

\clearpage

## COMMERCIAL BOTTOM TRAWL CATCH PER UNIT EFFORT INDICES

(ref:trawl-cpue-cap) Example commercial bottom trawl catch per unit effort (CPUE)
trends for `r sp`. Solid lines represent CPUE trends
standardized with a Tweedie GLMM (generalized linear mixed effects model) for
depth, latitude, DFO locality region, vessel, and month of year (Appendix
\@ref(app:cpue-models)). The line itself represents the estimate and the shaded
ribbon represents a 95\% confidence interval. The dashed line represents an
unstandardized commercial CPUE index calculated as the sum of catch divided by
the sum of effort each year. Standardized time series are scaled to have the
same maximum 95\% confidence interval. Unstandardized time series are scaled to
have the same geometric mean as the standardized time series. These are relative
index values --- the absolute value of the time series is not particularly
useful because it depends on arbitrary levels that the standardization
variables are set to. Management areas, as indicated in the top left corner of
each panel, are shown in Figure \@ref(fig:management-map).

```{r trawl-cpue-index, fig.asp=1.5, fig.width=4.5, out.width="3.5in", warning=FALSE, fig.cap="(ref:trawl-cpue-cap)"}
g$cpue_index
```

\clearpage

## MAPS OF COMMERCIAL CATCH PER UNIT EFFORT

(ref:cpue-maps-cap) Example commercial trawl and commercial hook-and-line
catch-per-unit-effort maps for Pacific Cod (note this figure is **not** `r sp`).
Lighter shading indicates higher levels
of a geometric mean of catch per unit effort in a given hexagonal cell. The
colour scale is fourth-root transformed to render a visual pattern similar to
a log transformation without overemphasizing differences close to zero. The
colour scale ('Viridis') is perceptually uniform, robust to colour blindness,
and prints accurately in grayscale [@viridis2018]. Cells are 7 km wide and are
only shown in cases where there are at least 3 unique vessels in a given cell
to meet privacy requirements. For bottom trawl, catch per unit effort is
calculated as the weight of catch (landings plus discards) divided by hours
fished for all positive tows from the groundfish trawl sector. Trawl data are
shown from 2013 onwards when the trawl footprint was frozen. Trawl data from 2007--2012
are indicated as outlined light grey hexagons to illustrate fishing
prior to the frozen footprint. For hook and line,
catch per unit effort is shown as the number of fish recorded as landed or
discarded per set. Hook-and-line data are shown from 2008 onwards.
Including as many years of data as possible reduces the number of discarded
fishing events when implementing the 3-vessel privacy requirement.
`r map_text`

```{r cpue-maps, fig.asp=1.37, fig.width=4.65, out.width="3.6in", fig.cap="(ref:cpue-maps-cap)"}
g_alt <- readRDS(here::here("report/ggplot-objects/pacific-cod.rds"))
trans <- "fourth_root_power"

p1_trawl <-
  g_alt$cpue_spatial +
  labs(fill = "Geometric\nmean of trawl\nCPUE (kg/hr)") +
  labs(colour = "Geometric\nmean of trawl\nCPUE (kg/hr)") +
  ggplot2::scale_fill_viridis_c(trans = trans, option = "D", breaks = c(2, 10, 50, 200)) +
  ggplot2::scale_colour_viridis_c(trans = trans, option = "D", breaks = c(2, 10, 50, 200)) +
  guides(fill = ggplot2::guide_colorbar(), colour = ggplot2::guide_colorbar()) +
    theme_pbs() + theme(legend.position = "bottom")

p2_ll <-
  g_alt$cpue_spatial_ll +
  labs(fill = "Geometric\nmean of hook-and-line\nfish/set") +
  labs(colour = "Geometric\nmean of hook-and-line\nfish/set") +
  ggplot2::scale_fill_viridis_c(trans = trans, option = "D", breaks = c(2, 5, 10, 20)) +
  ggplot2::scale_colour_viridis_c(trans = trans, option = "D", breaks = c(2, 5, 10, 20)) +
  guides(fill = ggplot2::guide_colorbar(), colour = ggplot2::guide_colorbar()) +
    theme_pbs() + theme(legend.position = "bottom")

gridExtra::grid.arrange(
  p1_trawl, # + checking_square,
  p2_ll,
  nrow = 1)
```

\clearpage

## AVAILABLE BIOLOGICAL SAMPLES {#sec:bio-samples}

```{r samples, warning=FALSE, fig.asp=0.5, fig.width=7, out.width="5.2in", fig.cap=paste0("Example specimen-availability plot for ", sp, ". Shown are the number of available fish specimens that have had their length measured, have been weighed, had their maturity assessed, had their age assessed, and for which ageing structures are available for ageing. Data are shown across all surveys (not just surveys shown elsewhere in the synopsis; top panel) and across all commercial fleets (bottom panel). Blank panels indicate year-measurement combinations without any data. Shading of these cells reflects the relative number of specimens available with the actual number of specimens indicated in the cells to the nearest round number.")}
gridExtra::grid.arrange(
  g$survey_samples,
  g$comm_samples,
  nrow = 2
)
```

\clearpage

## LENGTH COMPOSITION DATA

```{r lengths, warning=FALSE, fig.asp=0.8, fig.cap=paste0("Example length-frequency plot for ", sp, ". Female fish are shown as coloured (or black) bars and male fish are shown behind as light grey bars. The total number of fish measured for a given survey and year are indicated in the top left corner of each panel. Histograms are only shown if there are more than 20 fish measured for a given survey-year combination. The commercial male and female fish are combined since many are unsexed. See Figure \\ref{fig:survey-index} for survey abbreviations.")}
g$lengths
```

\clearpage

## AGE COMPOSITION DATA

(ref:ages-cap) Example age-frequency plot for `r sp`. Female fish are shown as
coloured (or black) circles and male fish are shown behind as light grey
circles. The total number of fish aged for a given survey and year are indicated
along the top of the panels. Diagonal lines are shown at five-year intervals to
facilitate tracing cohorts through time. See Figure \@ref(fig:survey-index) for
survey abbreviations. Ageing precision plots comparing precision of readings by
two individuals ageing the fish are provided for all species for which age data
exist in Appendix \@ref(app:age-precision).

```{r ages, fig.width=9, fig.asp=0.68, warning=FALSE, fig.cap="(ref:ages-cap)"}
g$ages
```

<!-- TODO: Implement of the weighted lengths and ages? If I do then adjust the caption accordingly and reference the original weighting. -->

\clearpage

## LENGTH-AGE AND LENGTH-WEIGHT MODEL FITS

\mathchardef\mhyphen="2D % math hyphen

```{r length-weight-vb, warning=FALSE, fig.asp=0.4, fig.cap=paste0("Example length-age and length-weight model fits and plots for ", sp, ". The length-age growth curve is a von Bertalanffy model of the form $L_i \\sim \\operatorname{Log-normal} \\left( \\log(l_\\mathrm{inf} (1 - \\exp(-k (A_i - t_0)))), \\sigma \\right)$ where $L_i$ and $A_i$ represent the length and age of fish $i$, $l_\\mathrm{inf}$, $k$, and $t_0$ represent the von Bertalanffy growth parameters, and $\\sigma$ represents the scale parameter. The length-weight curve is of the form $\\log (W_i) \\sim \\mathrm{Student\\mhyphen t} (df = 3, \\log (\\ln(a) + b L_i), \\sigma),$ with $W_i$ and $L_i$ representing the weight and length for fish $i$ and $\\sigma$ representing the observation error scale. We set the degrees of freedom of the Student-t distribution to 3 to be robust to outliers. The variables $a$ and $b$ represent the estimated length-weight parameters. Female model fits are indicated as solid black lines and male model fits are indicated as dashed grey lines. Text on the panels shows the parameter estimates and open grey circles represent individual fish that the models are fit to. These figures include all survey samples. See Appendices \\ref{sec:length-age-models} and \\ref{sec:length-weight-models} for details on the models.")}
cowplot::plot_grid(g$vb, g$length_weight, align = "v")
```

<!-- % TODO: I only fit models in cases where there were more than 100 fish samples for -->
<!-- % a given species-sex combination. -->

\clearpage

## MATURITY FREQUENCY BY MONTH

```{r maturity-months, out.width="4in", fig.asp=0.55, fig.width=5, fig.cap=paste0("Example maturity-frequency-by-month plot for ", sp, ". Categories of maturity are listed from most immature (top) to most mature (bottom); individual fish, once mature, cycle through the mature stages. The area of each circle corresponds to the number of fish specimens in a given maturity category for the given month. Female fish are indicated by black circles and male fish are indicated by light grey circles behind. The total number of fish specimens for each month are indicated by the numbers at the top of the plot. This plot includes data from both the commercial and survey samples.")}
g$maturity_month
```

## MATURITY OGIVES

```{r maturity-ogives, fig.asp=0.37, fig.cap=paste0("Example age- and length-at-maturity ogive plots for ", sp, ". Maturity ogives are fit as logistic regressions to individual fish specimens, which are categorized as mature vs. not mature. The solid black lines represent fits to the female fish and the dashed grey lines represent fits to the male fish. The vertical lines indicate the estimated age or length at 50\\% maturity. Text on the panels indicates the estimated age and length at 5, 50 and 95\\% maturity for females (F) and males (M). Model fits are only shown for cases where there are at least 20 mature and 20 immature males and females. Short rug lines along the top and bottom of each panel represent up to 1500 randomly chosen individual fish with a small amount of random jittering in the case of ages to help differentiate individual fish. Models are fit to all available survey samples regardless of time of year. See Appendix \\ref{sec:maturity-models} for details.")}
cowplot::plot_grid(g$mat_age, g$mat_length, align = "v")
```

\clearpage
