\clearpage

# ACKNOWLEDGEMENTS

We are grateful for the ongoing collaboration between DFO, commercial fishers,
First Nations, and non-governmental organizations, which has made possible the
collection of the valuable data that underlies this report. We thank
the project's steering committee (Greg Workman, Robyn Forrest, Dana Haggarty,
Chris Grandin, and Rob Kronlund) for invaluable input into the report design
and feedback throughout its production.
We thank Chris Rooper and Daniel Ricard for their thorough and helpful reviews.
We thank Mary Thiess and Kieran Forge for chairing and rapporteuring the review meeting, respectively.
We thank Christie Whelan for her
support initiating this project. We thank Norm Olsen, Maria Surry, and Malcolm
Wyeth for providing support on data extraction and general database structure
and content. We thank the participants of the peer review meeting on tiered
approaches in May 2016 [@macdougall2016], which included a data scorecard by
Norm Olsen, from which this report takes inspiration. We also thank Norm Olsen
for his work on GFASTeR, from which this project takes inspiration and
borrows some data queries. We thank Rowan Haigh for his work maintaining
the PBSmapping and PBSdata R packages, and indeed the authors of all the R
packages necessary to build this report (Appendix \@ref(app:reproducibility)).

\clearpage

`r if(knitr:::is_latex_output()) '\\Appendices'`

`r if(!knitr:::is_latex_output()) '# (APPENDIX) Appendix {-}'`

# AGEING PRECISION {#app:age-precision}

```{r age-precision-make-data, warning=FALSE, message=FALSE}
dc <- here::here("report", "data-cache")
spp <- gfsynopsis:::get_spp_names() %>% filter(type == "A")
spp <- filter(spp, species_common_name != "pacific hake")

out_dat <- list()
plots <- list()
for (i in seq_along(spp$species_common_name)) {
  dat <- readRDS(paste0(file.path(dc, spp$spp_w_hyphens[i]), ".rds"))
  if (nrow(dat$age_precision) > 1) {
    out_dat[[i]] <- gfplot::tidy_age_precision(dat$age_precision)
    out_dat[[i]]$species_common_name <-
      gfsynopsis:::first_cap(spp$species_common_name[[i]])
    plots[[i]] <- out_dat[[i]] %>%
      gfplot::plot_age_precision() +
      ggplot2::ggtitle(gfsynopsis:::first_cap(spp$species_common_name[[i]]))
  }
}
```

```{r age-precision-grid, warning=FALSE, message=FALSE, fig.width=8, fig.asp=1.25, out.width="5.0in", fig.cap="Ageing precision plots for all species in the report with data. Each dot and cross-hatch represents an individual fish that has been aged twice. The x-axis represents the age and upper and lower possible ages recorded by the initial ('primary') individual ageing the fish. The y-axis represents the equivalent values recorded by the second ('precision') individual ageing the fish. The dashed diagonal line represents a perfect one-to-one agreement between the two ages. Up to 250 fish have been randomly sampled from all fish precision-aged for a species, and a small amount of random jitter has been added to both axes to reduce overplotting with the same jitter value added to both the x and y axes for a given fish."}
out_prec <- bind_rows(out_dat)
set.seed(42)
jitter <- 0.2
jit <- stats::runif(nrow(out_prec), -jitter, jitter)
out_prec$prec_age <- out_prec$prec_age + jit
out_prec$prim_age <- out_prec$prim_age + jit
out_prec$prim_min_age <- out_prec$prim_min_age + jit
out_prec$prim_max_age <- out_prec$prim_max_age + jit
out_prec$prec_min_age <- out_prec$prec_min_age + jit
out_prec$prec_max_age <- out_prec$prec_max_age + jit

out_prec %>% group_by(species_common_name) %>%
  do(if (nrow(.) > 300) sample_n(., 300) else .) %>% # downsample for clarity
  # fake data points to get 1-1 aspect ratio (they are invisible; coloured NA):
  do(data.frame(
    type = c(rep("real", length(.$prim_max_age)), "fake"),
    prim_max_age = c(.$prim_max_age, max(c(.$prim_max_age, .$prec_max_age))),
    prec_max_age = c(.$prec_max_age, max(c(.$prim_max_age, .$prec_max_age))),
    prim_min_age = c(.$prim_min_age, 0),
    prec_min_age = c(.$prec_min_age, 0),
    prec_age = c(.$prec_age, 0),
    prim_age = c(.$prim_age, 0),
    species_code = c(.$species_code, unique(.$species_code)),
    species_common_name = c(.$species_common_name, unique(.$species_common_name)),
    stringsAsFactors = FALSE
    )) %>%
  ungroup() %>%
  mutate(species_common_name = gsub("Rougheye\\/blackspotted Rockfish Complex",
    "Rougheye/\nblackspotted Rockfish", species_common_name)) %>%
  mutate(species_common_name =
      forcats::fct_reorder(species_common_name, as.numeric(as.factor(species_code)))) %>%
  ggplot(aes_string("prim_age", "prec_age", colour = "type")) +
  geom_point(pch = 21, alpha = 0.6) +
  ggplot2::geom_abline(intercept = 0, slope = 1, col = "grey50", lty = 2) +
  ggplot2::geom_segment(aes_string(
    x = "prim_min_age", xend = "prim_max_age",
    y = "prec_age", yend = "prec_age"),
    alpha = 0.4) +
  facet_wrap(~species_common_name, scales = "free") +
  ggplot2::geom_segment(aes_string(
    x = "prim_age", xend = "prim_age",
    y = "prec_min_age", yend = "prec_max_age"), alpha = 0.4) +
  labs(title = "Ageing precision", x = en2fr("Primary age (years)", french),
    y = en2fr("Secondary age (years)", french)) +
  theme_pbs() +
  ggplot2::ggtitle("") +
  scale_colour_manual(values = c("real" = "grey25", "fake" = NA)) +
  guides(colour = FALSE)
```
