\vspace{0mm}

\vspace{0mm}

# DEPTH-DENSITY RELATIONSHIPS FOR THE SYNOPTIC SURVEYS {#app:depth-curves}

```{r sdmTMB-depth-all-calculations}
surveys <- c("SYN QCS", "SYN HS", "SYN WCHG", "SYN WCVI")
fi <- list.files(here::here("report/map-cache/synoptic"), full.names = FALSE)
out <- purrr::map_df(fi, function(i) {
  mm <- readRDS(here::here(paste0("report/map-cache/synoptic/", i)))
  out <- purrr::map_df(1:4, function(ii) {
    if (length(mm$models[[ii]]$models) > 1L) {
      rd <- dplyr::filter(mm$raw_dat, survey == surveys[ii])
      # x <- seq(min(rd$depth_scaled, na.rm = TRUE),
      #   max(rd$depth_scaled, na.rm = TRUE), length.out = 200)
      range_d <- -1 * exp(rd$depth_mean[1] +
          range(rd$depth_scaled, na.rm = TRUE) * rd$depth_sd[1])
      x <- seq(-3, 4, length.out = 300)
      x2 <- x^2
      B <- mm$models[[ii]]$models$model$par
      y <- B[[1]] + x * B[[2]] + x2 * B[[3]]
      out <- data.frame(
        depth = -1*exp(x * rd$depth_sd[1] + rd$depth_mean[1]),
        y = exp(y) * 1000, # convert to kg/km^2 (was scaled by 1000 already)
        survey = surveys[ii],
        species = gsub("-", " ", gsub(".rds", "", i)),
        stringsAsFactors = FALSE)
      out <- mutate(out,
        extrapolated = depth < min(range_d) * 1 |
          depth > max(range_d) * 1)
      out
    }
  })
  out
})

dd <- out %>% group_by(species, survey) %>%
  # mutate(y = y / max(y))
  mutate(max_y = max(y[!extrapolated])) %>%
  mutate(y = ifelse(y < max_y * 1.15, y, NA)) %>%
  mutate(mode_depth = depth[y == max(y)[1]]) %>%
  group_by(species) %>%
  mutate(mean_mode_depth = mean(mode_depth)) %>%
  filter(depth >= -800)

make_depth_plot <- function(.data) {
  ggplot(.data, aes_string('depth', 'y', colour = 'survey')) +
    geom_line(lty = 2) +
    coord_cartesian(xlim = c(-800, 0), ylim = ) +
    labs(x = 'Depth (m)',
      y = expression(Survey~biomass~density~(kg/km^2)), colour = 'Survey') +
    # facet_wrap(~forcats::fct_reorder(species,
    #   mean_mode_depth),
    facet_wrap(~gfsynopsis:::first_cap(species), scales = "free_y", ncol = 4) +
    geom_line(data = dplyr::filter(.data, !extrapolated), lwd = 0.9) +
    scale_color_brewer(palette = "Dark2")
}
```

(ref:depth-relationship1) Predicted relationships between depth and biomass
density for all species across the four synoptic surveys (Part 1 of 2). Solid
lines indicate relationships predicted within the depth range of the survey and
dashed lines indicate extrapolated relationships beyond the observed depth.
These relationships are derived from the depth coefficients in the spatial
models that generate the map plots (e.g. Figure \@ref(fig:survey-maps)). These
plots provide some indication about which surveys encompass the entirety of the
depth distribution for particular species. See Appendix
\@ref(app:spatial-modeling) for details on the models underlying these
predictions.

```{r sdmTMB-depth-all-plots1, fig.width=10, fig.asp=1.10, out.width="\\textwidth", fig.cap="(ref:depth-relationship1)", fig.pos="H"}
.sp <- sort(unique(dd$species))
filter(dd, species %in% .sp[1:36]) %>%
  make_depth_plot()
```

\clearpage
\vspace{0mm}

(ref:depth-relationship2) Part 2 of Figure \@ref(fig:sdmTMB-depth-all-plots1). Caption is the same otherwise.

```{r sdmTMB-depth-all-plots2, fig.width=10, fig.asp=1.3, out.width="\\textwidth", fig.cap="(ref:depth-relationship2)", fig.pos="p"}
filter(dd, species %in% .sp[37:length(.sp)]) %>%
  make_depth_plot()
```

\clearpage

\vspace{0mm}

\vspace{0mm}
