# SPATIAL MODELLING OF SURVEY BIOMASS {#app:spatial-modeling}

We modelled the expected biomass density in space for each species using
geostatistical models applied to data from the fisheries independent bottom
trawl and long line surveys (e.g. Figure \@ref(fig:survey-maps)). Our modeling
approach is consistent with recent models used for spatial modelling and
spatiotemporal index standardization of groundfish populations [e.g.
@shelton2014; @ward2015; @thorson2015; @thorson2016]. Such models have been
shown, for example, to improve estimates of rockfish abundance and distribution
[@shelton2014] and improve precision when estimating relative abundance indices
for groundfish [@thorson2015]. Our specific model is fit with TMB
[@kristensen2016] in R [@r2018] with the help of INLA [@rue2009; @lindgren2015]
via the R package sdmTMB, which we wrote for this purpose (Appendix
\@ref(app:reproducibility)).

At a high level, these models predict relative biomass or catch rate in space
as a continuous process with a quadratic effect for bottom depth, spatial
random effects that represent an amalgamation of spatial processes not
explicitly included in the model, and an observation error component. After
fitting the model to survey sets from trawl or longline surveys, we then
project the model predictions onto a 2 $\times$ 2 km grid in a UTM 9 projection
to derive estimates of biomass throughout the survey domain.

Similarly to the commercial catch per unit effort standardization models
(Appendix \@ref(app:cpue-models)), these models can be represented as Tweedie
GLMMs with a log link:

\begin{align}
  y_s &\sim \mathrm{Tweedie}(\mu_s, p, \phi), \quad 1 < p < 2,\\
  \mu_s &= \exp \left( \bm{X}_s \bm{\beta} + \omega_s \right),
\end{align}

where $s$ represents a spatial location, $y_s$ represents observed fish density
for a survey set, $\mu_s$ represents expected fish density, $p$ represents the
Tweedie power parameter, and $\phi$ represents the Tweedie dispersion
parameter. The symbol $\bm{X_s}$ represents a vector of predictors (an
intercept, log depth, and log depth squared) and $\bm{\beta}$ represents
a corresponding vector of coefficients. The spatial random effects $\omega_s$
are assumed to be drawn from a multivariate normal distribution with
a covariance matrix $\bm{\Sigma}_\omega$ that is centered on zero:

$$\bm{\omega} \sim \mathrm{MVNormal} \left( \bm{0}, \bm{\Sigma}_\omega \right).$$

We constrain the spatial random effects to follow a \mbox{Mat\'ern} covariance
function, which defines the rate with which spatial correlation decays with
distance (Figure \@ref(fig:sdmTMB-matern-demo)). The \mbox{Mat\'ern} function
describes the covariance $\Phi \left( s_j, s_k \right)$ between spatial
locations $s_j$ and $s_k$ as:

$$\Phi\left( s_j,s_k \right) = \tau^2/\Gamma(\nu)2^{\nu - 1}
    (\kappa d_{jk})^\nu K_\nu \left( \kappa d_{jk} \right),$$

where $\tau^2$ represents the spatial variance, $\Gamma$ represents the Gamma
function, $K_\nu$ represents the Bessel function, $d_{jk}$ represents the
Euclidean distance between locations $s_j$ and $s_k$, and $\kappa$ represents
a scaling parameter that is estimated [e.g. @lindgren2011]. The parameter $\nu$
controls the smoothness of the covariance function. We set $\nu = 1$, which
lets us take advantage of the Stochastic Partial Differential Equation (SPDE)
approximation to Gaussian Markov Random Fields (GMRF) to greatly increase
computational efficiency [@lindgren2011].

Our spatial model falls into the general category of "predictive process"
models [e.g. @latimer2009; @shelton2014; @anderson2018], in which the model
keeps track of a limited number of "knots" that approximate unexplained spatial
variation (Figure \@ref(fig:sdmTMB-spde)). By keeping track of a smaller number
of knots than the full spatial data set, we can increase computational
efficiency. The model predictions can be projected to the original data
locations or any new set of locations as long as an appropriate covariance
matrix can be calculated [e.g. @latimer2009]. Higher numbers of knots result in
a better approximation of the spatial random effects at a greater computational
cost. We fit the spatial models with one less knot than the number of
observations if there were fewer than 200 observations and 200 knots otherwise.
Following one common practice [e.g. @shelton2014; @thorson2017b; @anderson2018]
we chose the location of the knots with a k-means clustering algorithm with
a fixed random seed to ensure reproducible results.

Instead of directly modelling the effects of depth and depth squared, we first
centred the log-transformed depth covariate by subtracting its mean and scaling
it by its standard deviation. We then calculate "depth squared" from this
centred and scaled variable. This ensures the covariate values are not too
large or small to avoid computational issues and the centering separates the
linear and quadratic predictor components.

We fit the four synoptic survey data sets separately, because only two of the
surveys are conducted each year and the surveys are disjointed in space and
time. Similarly, we fit the North and South HBLL surveys independently. We
combined the predictions to generate the map plots, but labelled the years in
which the various surveys were conducted.

As an example, we illustrate the model components for Pacific Cod in Queen
Charlotte Sound (Figure \@ref(fig:sdmTMB-maps-combined)). We begin with
a bathymetry layer and biomass density value for each survey set in space
(Figure \@ref(fig:sdmTMB-maps-combined)A). After fitting the model, we can
inspect the effect of the bottom depth quadratic fixed effect predictors
(Figure \@ref(fig:sdmTMB-maps-combined)B) as well as the spatial random effects
(Figure \@ref(fig:sdmTMB-maps-combined)C). If we add the fixed effect
predictions to the spatial random effects in link (log) space and exponentiate
the result we derive model predictions that include both the fixed and random
effects (Figure \@ref(fig:sdmTMB-maps-combined)D). We can inspect randomized
quantile residuals in space to check for any remaining spatial correlation
(Figure \@ref(fig:sdmTMB-resids)). We can also look at the predicted
relationship between depth and biomass density across all the species (Figures
\@ref(fig:sdmTMB-depth-all-plots1), \@ref(fig:sdmTMB-depth-all-plots2)).

```{r sdmTMB-matern-demo, fig.asp=0.3, fig.width=8, fig.cap="An illustration of the effect of the $\\kappa$ parameter ('kappa') on the shape of the \\mbox{Mat\\'ern} correlation function with $\\nu = 1$. The vertical dashed line illustrates $\\sqrt{8\\nu} / \\kappa$, referred to as the 'range', which is a point at which the correlation decreases below 0.1."}
library(dplyr)
library(ggplot2)
dist <- seq(0, 13, length = 200)
kappa <- c(0.3, 0.5, 1.1)
nu <- 1.0
plyr::ldply(kappa, function(k) {
  data.frame(
    Correlation = k * dist * base::besselK(k * dist, nu),
    kappa = paste("kappa =", k),
    Distance = dist,
    range = sqrt(8 * 1) / k
  )
}) %>%
  mutate(Correlation = ifelse(is.na(Correlation), 1, Correlation)) %>%
  ggplot(aes(Distance, Correlation)) +
  geom_line() +
  facet_wrap(~kappa) +
  geom_vline(aes(xintercept = range), lty = 2) +
  coord_cartesian(expand = FALSE, ylim = c(0, 1)) +
  theme(strip.text.x = element_text(colour = "grey20", size = rel(1.1)))
```

```{r sdmTMB-data}
library(sdmTMB)
d <- readRDS(here::here("report/data-cache/pacific-cod.rds"))
d <- d$survey_sets
dat <- gfplot:::tidy_survey_sets(d, "SYN QCS", years = 2017)
dat <- mutate(dat, density = density*1000*1000)
dat <- filter(dat, !is.na(depth))
dat <- gfplot:::scale_survey_predictors(dat)
dat <- select(dat, -X10, -Y10)

grid_locs <- gfplot:::make_prediction_grid(filter(dat, year == 2017),
  survey = "SYN QCS", cell_width = 1.5)$grid
grid_locs <- rename(grid_locs, depth = akima_depth)
grid_locs$year <- NULL
```

```{r sdmTMB-spde, fig.width=9, out.width="3.5in", fig.cap="Example triangularization mesh for Pacific Cod and the Queen Charlotte Sound survey in 2017. The red dots indicate knot locations. The open grey circles in the background represent the locations of the observed data. The lines show the triangularization mesh used in the SPDE approximation. In this case, many of the knots overlap the observed data. A greater number of knots will increase the accuracy of the approximation at the expense of computational time."}
spde <- make_spde(dat$X, dat$Y, n_knots = 200)
plot_spde(spde)
```

```{r sdmTMB-model}
m <- sdmTMB(
  data = dat, formula = density ~ depth_scaled + depth_scaled2,
  time = "year", spde = spde, family = tweedie(link = "log"),
  silent = TRUE, anisotropy = FALSE
)
```

```{r sdmTMB-preds, fig.width=6, out.width="4in"}
predictions <- predict(m, newdata = grid_locs)
plot_map <- function(.dat, column) {
  ggplot(.dat, aes_string("X", "Y", fill = column)) +
    geom_raster() +
    coord_fixed(expand = FALSE) +
    xlab("Easting") + ylab("Northing") +
    theme(legend.position = "bottom") +
    geom_point(data = dat, aes(X, Y, size = density), pch = 21, inherit.aes = FALSE) +
    scale_size_area() +
    guides(size = FALSE)
}
```

```{r sdmTMB-plot-depth}
g_depth <- plot_map(grid_locs, "depth") +
  scale_fill_distiller(trans = "fourth_root_power", palette = "Blues", direction = 1,
    breaks = c(50, 200, 600)) +
  ggtitle("(A) Depth") +
  labs(fill = 'Depth (m)')
```

```{r sdmTMB-plot-combined}
limits <- c(0, max(exp(predictions$data$est))*1.001)
breaks <- c(1, 20, 150)
g_sdm1 <- plot_map(predictions$data, "exp(est)") +
  scale_fill_viridis_c(trans = "fourth_root_power", option = "C",
    breaks = breaks, limits = limits) +
  ggtitle("(D) Prediction (depth effects + spatial random effects)") +
  labs(fill = 'Biomass estimate\nkg/km^2')
```

```{r sdmTMB-plot-fe}
g_sdm2 <- plot_map(predictions$data, "exp(est_fe)") +
  ggtitle("(B) Prediction (depth effects only)") +
  scale_fill_viridis_c(trans = "fourth_root_power", option = "C",
    breaks = breaks, limits = limits) +
  labs(fill = 'Biomass estimate\nkg/km^2')
```

```{r sdmTMB-plot-re}
g_sdm3 <- plot_map(predictions$data, "est_re_s") +
  ggtitle("(C) Spatial random effects") +
  scale_fill_gradient2(midpoint = 0,
    high = scales::muted("red"),
    mid = "white",
    low = scales::muted("blue")) +
  labs(fill = 'Deviation from expected\nbiomass in log space')
```

\clearpage

```{r sdmTMB-maps-combined, fig.width=9, fig.asp=1, out.width="6in", fig.cap="Example spatial model components in Queen Charlotte Sound for Pacific Cod. (A) Bottom depth data that is used as a predictor. (B) Predicted biomass density from the quadratic effect of depth only. (C) Spatial random effect deviations. These deviations represent on modelled correlated spatial effects. (D) Model predictions that include both the depth fixed effects and the spatial random effects. Circles represent the biomass density for each survey set with the area of the circle proportional to the density."}
cowplot::plot_grid(g_depth, g_sdm2, g_sdm3, g_sdm1)
```

\clearpage

```{r sdmTMB-resids, fig.width=6, out.width="4in", fig.cap="Spatial randomized quantile residuals on the link (log) scale."}
dat$resids <- residuals(m) # randomized quantile residuals
# hist(dat$resids)
# qqnorm(dat$resids);abline(a = 0, b = 1)
ggplot(dat, aes(X, Y, col = resids)) + scale_colour_gradient2() +
  geom_point() + coord_fixed() +
  xlab("Easting") + ylab("Northing") +
  scale_colour_gradient2(midpoint = 0,
    high = scales::muted("red"),
    mid = "white",
    low = scales::muted("blue")) +
  labs(colour = 'Residual')
```
