# CPUE INDEX STANDARDIZATION {#sec:cpue-models}

We present two commercial bottom trawl catch per unit effort (CPUE) indices in
the synopsis report: a 'raw' unstandardized timeseries and a 'standardized' time
series (e.g. Figure \@ref(fig:trawl-cpue-index). The standardized time series
follows the approach commonly employed for British Columbia groundfish stock
assessments of using a Generalized Linear Model (GLM) to account for the
confounding effects of month, depth, latitude, vessel, and DFO locality
(predesignated historical geographical locations) when calculating an annual
index value. We do this by fitting two GLMs: one to data representing whether or
not the species of interest (e.g. Pacific Cod for a Pacific Cod CPUE index) is
caught in a given tow (the 'binary' model),
and a second model to data representing the CPUE conditional on a tow having caught
at least one of that species of interest (the 'lognormal' model). This approach is
sometimes called a delta-lognormal GLM. These two model components can then be
combined to derive the final estimate of CPUE for a given year at a consistent
value for all the predictors, i.e. a consistent month, depth, latitude, vessel,
and DFO locality.

For depth and latitude, we binned the values into a sequence of bands to allow
for nonlinear relationships between these predictors and the response. For
latitude, we used bands that are 0.2 degrees wide, start at the 2% quantile of
the latitude values for the specific region and end at the 98% quantile of the
latitude values. We used this quantile method to algorithmically account for
outlying values that were often data recording errors and could cause problems
when fitting the models. We rounded the lowest latitude bin down to the nearest
0.2 latitude value and rounded to the upper latitude bin up to the nearest 0.2
latitude value. For depth, we used the fishing capture depth binned into bands
that are 50m wide. The bins start at the 1% quantile of the depth values for
the specific region and end at the 99% quantile of the depth values. We rounded
the lowest and highest bin values to 50m values similarly to the process for
the latitude bins as described above.

Due to the multispecies nature of the BC groundfish fishery, it is necessary to
define rules about which vessels should be considered part of a 'fleet'
with which to calculate a CPUE index. We follow the approach used in a number of
recent BC groundfish stock assessments by requiring vessels to have caught the
species in a certain number of tows across all years of interest, and to have
passed a certain threshold of positive trips (trips that recorded some of the
species) for a certain number of years. Our current implementation requires a
vessel to have recorded at least 100 positive tows since 1996 and to have
recorded at least four positive trips in at least four years since 1996. These
decisions are in line with recent BC groundfish stock assessments
[e.g. @starr2017pollock].

We fit the binomial model (denoted with a superscript $B$) as a logistic
regression:

\begin{align}
  y_i^B &\sim \mathrm{Binomial}(\pi_i^B)\\
  \mathrm{logit}\left(\pi_i^B\right) &= \bm{X}_i^B \bm{\beta}^B,
\end{align}

and $i$ represents a single tow, $y_i$ represents either a 1 if a tow caught the
species or a 0 if it did not, $\bm{X^B_i}$ represents a vector of predictors,
$\bm{\beta^B}$ represents a vector of coefficients, and $\pi_i^B$ represents the
estimated probability of observing the species in a tow. Details to follow on
the specific depth and latitude bands chosen.

We fit the lognormal model (denoted with a superscript $L$) as a linear model
fit to log-transformed response data:

\begin{equation}
  \log \left(y_i^L\right) \sim \mathrm{Normal} \left(\bm{X}_i^L \bm{\beta}^L, \sigma \right),
\end{equation}

where the symbols can be interpreted as before except that $y_i^L$ represents
the CPUE in kg per hour towed for a tow that did catch at least one of the
species, and $\sigma$ represents the standard deviation of unexplained residual
error.

We can then calculate the standardized estimate of CPUE for year $t$, $\mu_t$ as:

\begin{equation}
  \mu_t = \mathrm{logit}^{-1} \left(\bm{X}_r^B
    \bm{\beta}^B \right) \cdot \exp \left(\bm{X}_r^L \bm{\beta}^L \right)
\end{equation}

or

\begin{equation}
  \mu_t = \pi_t^B \cdot \exp \left(\bm{X}_r^L \bm{\beta}^L \right)
\end{equation}

where $\bm{X_r}$ represents a vector of predictors set to the reference ($r$)
levels with the year set to the year of interest. We chose the reference levels
as the most frequent level of each predictor in the positive-only data
[@maunder2004]. For example, we set the reference month as the most common
month observed in the dataset filtered for only tows where the species was
caught. This will have a minor effect on the shape of the final CPUE index
because of the multiplication of the binary and positive components.

Recent BC groundfish stock assessments have used a bootstrap approach to derive
estimates of uncertainty around standardized CPUE timeseries. This proved to be
computationally unfeasible to run for all the management areas and stocks in
this synopsis report. Therefore we wrote the model in the statistical software
TMB [@kristensen2016] and use standard errors ($\mathrm{SE}$) as calculated
by TMB on $\log (\mu_t)$ via the Delta method as is commonly done for such
models [e.g. @thorson2015]. We then calculated the 95% confidence
intervals as $\exp (\mu_t \pm 1.96 \mathrm{SE}_t)$.

We calculated the 'raw' unstandardized timeseries is calculated using a similar
procedure but without any of the covariates other than a factor predictor for
each year. This is equivalent to calculating the geometric mean of CPUE each
year.

We tested our implementation of this index standardization model against
simulated data to ensure that (1) it can generate unbiased annual estimates and
(2) the confidence intervals have the correct coverage, i.e. the confidence
intervals contain the true known value at the expected frequency of 95%. We
also compared our estimates qualitatively to recent estimates from similar
models in published assessments to ensure we derived similar timeseries.

For the simulation testing, as examples, we present results for two simulations
with 10 vessels, each vessel recording data from 15 tows, 20 years of data, and
lognormal observation error with standard deviation 0.35 (i.e., approximately
a CV of 0.35). The standardization model then accounts for vessel ID in both
a binary and positive component model. The delta model derives an unconditional
estimate of CPUE after combining the two component models. The parameter
estimates from the component models regularly fall along the one-to-one line of
unbiased and accurate prediction (e.g. Fig. \@ref(fig:cpue-sim-cross-val)) and
the unconditional estimates of CPUE and their confidence intervals regularly
include the true values with correct coverage (e.g.
Fig. \@ref(fig:cpue-sim-ts-val)).

```{r cpue-model-load}
cpue_model <- readRDS(here::here("report/cpue-cache/petrale-sole-5AB-model.rds"))
```

```{r cpue-fleet-load}
# d_cpue <- readRDS(here::here("report/data-cache/cpue-index-dat.rds"))
d_cpue_pred <- gfplot::predict_cpue_index_tweedie(cpue_model, center = FALSE)
d_cpue_pred$formula_version <- "Full standardization"
if (file.exists(here::here("report/cpue-cache/petrale-sole-5AB-fleet.rds")))
  d_fleet <- readRDS(here::here("report/cpue-cache/petrale-sole-5AB-fleet.rds"))
```

```{r cpue-int-lines, fig.width=6, out.width="4in", fig.cap="Locality-specific CPUE index trends for a standardization model that allows for locality-year (space-time) interactions. The coloured lines indicate the locality-specific estimates with all other predictors set to their base levels. The black line and shaded ribbon indicate the overall average annual CPUE."}
if (exists("d_fleet"))
  plot_cpue_spaghetti(
    model = cpue_model,
    fleet = d_fleet,
    index_data = d_cpue_pred,
    era = "modern")
```

```{r cpue-tweedie-ex, fig.asp=0.7, fig.cap="Example density functions for the Tweedie distribution. The symbol $\\phi$ (written as phi in this figure) represents the dispersion parameter, $p$ represents the power parameter, and $\\mu$ represents the mean. Note that the spike in density that is seen towards the left of the panels is at a value of 0 on the x axis."}
plot_tweedie_ex <- function(df, max = 15) {
  xx <- seq(0, max, length.out = 1000)
  out <- plyr::mdply(df, function(mu, power, phi) {
    data.frame(x = xx,
      y = tweedie::dtweedie(y = xx, mu = mu, power = power, phi = phi))
  }) %>%
    mutate(phi = paste("phi =", phi)) %>%
    mutate(power = paste("p =", power)) %>%
    mutate(mu = paste("μ =", mu))

  ggplot(out, aes(x, y, colour = mu)) +
    geom_line() +
    facet_grid(power~phi, scales = "free") +
    gfplot::theme_pbs() +
    labs(colour = "μ") +
    xlab("Value") + ylab("Density") +
    coord_cartesian(expand = FALSE, xlim = c(-0.2, max(out$x))) +
    scale_colour_brewer(palette = "Dark2") +
    scale_fill_brewer(palette = "Dark2")
}

df <- expand.grid(power = c(1.2, 1.6, 1.8), mu = c(1, 3, 6), phi = c(0.5, 1, 2))
plot_tweedie_ex(df)
```

\clearpage

(ref:caption-cpue-quantile-residuals) Histograms of randomized quantile residuals [@dunn1996] for the CPUE GLMM standardization models. The histograms illustrate the actual density distribution of 10,000 randomly selected randomized quantile residuals. The dashed lines show the probability density for a normal distribution with the same standard deviation.

```{r cpue-quantile-residuals, fig.cap="(ref:caption-cpue-quantile-residuals)"}
qr <- gfplot::qres_tweedie(cpue_model, response = "cpue")
gfplot::plot_qres_histogram(qr)
```

```{r, fig.cap="3CD eg.", eval=FALSE}
readRDS(here::here("report/cpue-cache/petrale-sole-3CD-model.rds")) %>%
  gfplot::qres_tweedie(response = "cpue") %>%
  gfplot::plot_qres_histogram()
```

\clearpage

