# SURVEY BIOMASS INDEX TRENDS {#app:survey-trend-models}

Fisheries and Oceans Canada conducts a number of fishery-independent
multispecies research surveys annually or biannually. These include four
synoptic bottom trawl surveys and two longline hook surveys
(Figure \@ref(fig:intro-maps)). The survey areas correspond roughly with Groundfish
Management areas (Figure \@ref(fig:management-map)) that fishery quotas are
allocated by British Columbia waters. For the synoptic trawl surveys, the West
Coast Vancouver Island survey corresponds roughly to management areas 3CD, the
Queen Charlotte Sound survey corresponds roughly to areas 5AB, the Hecate
Strait survey corresponds roughly to areas 5CD, and the West Coast Haida Gwaii
surveys corresponds roughly to area 5E. The Hard Bottom Longline survey is
split into northern and southern segments. The southern survey area corresponds
roughly to management areas 3CD and 5AB while the northern survey area
corresponds roughly to areas\ 5CDE.

## SYNOPTIC BOTTOM TRAWL SURVEYS

DFO, together with the Canadian Groundfish Research and Conservation Society,
implemented a coordinated set of bottom trawl surveys that together cover the
continental shelf and upper slope of most of the BC coast. The surveys follow
a random depth stratified design and use the same bottom trawl fishing gear and
fishing protocols [@sinclair2003syn]. The surveys were designed to provide
a synopsis of all species available to bottom trawl gear as opposed to focusing
on specific species. There are a total of four synoptic (SYN) surveys: Hecate Strait
(HS), West Coast Vancouver Island (WCVI), Queen Charlotte Sound (QCS), and West
Coast Haida Gwaii (WCHG) (Figure \@ref(fig:intro-maps)). Surveys have been
conducted on both chartered commercial vessels and government research vessels.
The Hecate Strait survey and West Coast Vancouver Island survey have been
conducted on the Canadian Coast Guard research trawler WE Ricker while the
Queen Charlotte Sound survey has been conducted on chartered commercial fishing
vessels. Two of the synoptic surveys are conducted each year on an alternating
basis so that each survey is conducted every two years.

## HARD BOTTOM LONGLINE SURVEYS

The Pacific Halibut Management Association, in consultation with DFO, initiated
a depth-stratified, random design research longline survey conducted with
chartered commercial fishing vessels in 2006. These are referred to as the Hard
Bottom Longline Outside surveys. The survey employs standardized
longline snap gear and fishing methods and alternates annually between the
northern and southern portions of BC. The survey is designed to provide catch
rates of all species and biological samples of rockfish from the outside
coastal waters of BC for stock assessment, alternating between northern and
southern areas.

Hard Bottom Longline Inside surveys are conducted within management area 4B.
These surveys were designed to provide fishery independent indices of
abundance together with biological samples to improve the assessment of
Yelloweye (*Sebastes ruberrimus*) and Quillback (*S. maliger*) Rockfish
for the 4B management region. They began in the northern Strait of Georgia
in Pacific Fishery Management areas 12 and 13 in 2003 and 2004, and now
alternate years to cover the northern (areas 12 and 13) and southern
(areas 14--20, 28, 29) portions of the inside waters. These surveys also employ
standardized longline snap gear and fishing methods.

## INTERNATIONAL PACIFIC HALIBUT COMMISSION FISHERY INDEPENDENT SURVEY

The International Pacific Halibut Commission’s (IPHC) fishery independent
survey is the longest times series of longline survey data in BC.
It provides distribution, biomass, age, growth and maturity
data that are used in the annual assessment of Pacific Halibut
(*Hippoglossus stenolepis*). In Appendix \@ref(app:iphc-survey-index) we
describe how we use data from the survey to
construct a consistent index of abundance for other species
over as long a time series as possible, despite the survey design changing
through the years.

## FURTHER DETAILS OF SURVEYS

Details on the design of the various surveys referenced in this report can be
found in the following documents:

1. Synoptic Survey, Queen Charlotte Sound (SYN QCS): @williams2018synqcs

1. Synoptic Survey, West Coast Vancouver Island (SYN WCVI): @nottingham2017synwcvi

1. Synoptic Survey, Hecate Strait (SYN HS): @wyeth2018synhs

1. Synoptic Survey, West Coast Haida Gwaii (SYN WCHG): @williams2018synwchg

1. Hard Bottom Longline Survey, Outside (HBLL OUT): @dfo2018ifmp

1. Hard Bottom Longline Survey, Inside (HBLL INS): @lochead2007irf

1. Hecate Strait Multispecies Assemblage Survey (MSA HS): @choromanski2004hsmsa

1. International Pacific Halibut Commission fishery independent survey (IPHC FISS): @flemming2012iphc

The main biomass index trends illustrated on the figure pages represent the
'design-based' index trends that have historically been used in Pacific
Biological Station groundfish stock assessments. We extracted the trawl and
longline survey relative biomass index trends from GFBio, which are generated
using the same approach that has been used in all recent BC groundfish stock
assessment reports. The code to perform the calculations was originally written
by Norm Olsen at Pacific Biological Station and is automatically applied to the
available survey data to populate the GFBio database. We have included the relevant
equations below for clarity. We also compare geostatistical model-based
estimates of biomass index trends for the trawl surveys (Section \@ref(sec:geostatistical)).

## TRAWL SURVEY DESIGN-BASED ESTIMATION {#sec:trawl-methods}

For all trawl surveys, and for a given species, we calculated the relative
biomass density $B$ in year $t$ as:

\begin{equation}
  B_t = \sum_{i = 1}^k C_{t,i} A_i
\end{equation}

where $C_{t,i}$ represents the mean CPUE in kg/km^2^ for the species in year
$t$ and stratum $i$, $A_i$ represents the area of stratum $i$ in km^2^, and $k$
represents the number of strata. We calculated the CPUE ($C_{t,i}$) for a given
species in year $t$ and stratum $i$ as:

\begin{equation}
  C_{t,i} = \frac{\sum_{j = 1}^{n_{t,i}} \left( W_{t,j} / D_{t,j} w_{t,j}\right)}{n_{t,i}}
\end{equation}

where $W_{t,j}$ represents the catch weight (kg) for the species in year $t$,
stratum $i$, and tow $j$; $D_{t,j}$ represents the distance travelled in km by
tow $j$ in year $y$; $w_{t,j}$ represents the net opening width in km for year
$y$ and tow $j$; and $n_{t,i}$ represents the number of tows in year $t$ and
stratum $i$.

## LONG-LINE SURVEY DESIGN-BASED ESTIMATION {#sec:ll-methods}

For the HBLL surveys, and for a given species, we calculated the relative
biomass density $B$ in year $t$ as:

\begin{equation}
  B_t = \sum_{i = 1}^k C_{t,i} A_i
\end{equation}

where $C_{t,i}$ represents the mean CPUE in parts (fish) per
km\textsuperscript{2} for the species in year $t$ and stratum $i$, $A_i$
represents the area of stratum $i$ in km^2^, and $k$ represents
the number of strata. We calculated the CPUE ($C_{t,i}$) for a given species in
year $y$ and stratum $i$ as:

\begin{equation}
  C_{t,i} = \frac{\sum_{j = 1}^{n_{t,i}} \left( N_{t,j} / H_{t,j} w_{t,j}\right)}{n_{t,i}}
\end{equation}

where $N_{t,j}$ represents the number of fish caught for the species in year
$t$, stratum $i$, and set $j$; $H_{t,j}$ represents the number of hooks
$\times$ the hook spacing in km in set $j$ in year $t$; $w_{t,j}$ represents an
arbitrary swept width of 30 feet or 0.009144 km for year $t$ and tow $j$; and
$n_{t,i}$ represents the number of sets in year $t$ and stratum $i$. The hook
spacing is 8 feet or 0.0024384 km for the inside and outside HBLL surveys.

Details on the design-based estimation of biomass density index for the IPHC
survey are shown in Appendix \@ref(app:iphc-survey-index).

## DESIGN-BASED-INDEX CONFIDENCE INTERVALS

We calculated bootstrap confidence intervals on $B_t$ by repeatedly calculating
$B_t$ given the above equations but each time re-sampling, with replacement,
from the available tows within each stratum. We drew 1000 bootstrap replicates
of $B_t$, $B_t^{\mathrm{rep}}$, and calculated 95% bias-corrected and adjusted (BCa) confidence intervals [@efron1987] on $B_t^{\mathrm{rep}}$.

## GEOSTATISTICAL SPATIOTEMPORAL BIOMASS INDEX TRENDS {#sec:geostatistical}

The above-described design-based estimates assume that average fish density is
the same throughout each survey stratum and that the only source of variance
between samples is sampling stochasticity itself [@petitgas2001]. However, we
know this is not true --- a substantial portion of fish density variation
within a stratum can be attributed to habitat being of better or poorer
suitability for a given fish and this suitability can be because of many
factors beyond depth alone, on which the strata are stratified [@shelton2014].
Furthermore, design-based estimates assume that nearby strata are statistically
independent from each other. However, we know that fish do not perceive their
habitat according to these exact strata boundaries and that ecological
processes in general tend to be spatially correlated with processes closer to
each other being more similar than those further apart.

Geostatistical modelling of survey data aims to address these issues by
modelling fish density as a smooth spatial surface --- possibly the result of
explicit habitat variables such as depth --- but also as the product of other
unobserved or 'latent' spatial effects. In recent years, there has been
a movement toward such 'model-based' standardization of survey biomass index
trends [e.g. @shelton2014; @thorson2015; @webster2017]. We include model-based
index trends for the synoptic trawl surveys as a point of comparison so the
reader can gauge when the two approaches may differ. Large differences are
likely a result of the random positioning of the survey sets for a given year
ending up in particularly good or poor habitat for a given species. Authors of
BC groundfish stock assessments may want to consider model-based index
standardization on a case-by-case basis.

We use the geostatistical model as described in Appendix
\@ref(app:spatial-modeling) with the addition of spatiotemporal random
effects ($\epsilon_{s,t}$; defined for locations in space $s$ and time $t$)
and annual predictors for the mean biomass each year:

\begin{align}
  y_{s,t} &\sim \mathrm{Tweedie}(\mu_{s,t}, p, \phi), \quad 1 < p < 2,\\
  \mu_{s,t} &= \exp \left( \bm{X}_{s,t} \bm{\beta} + \omega_s + \epsilon_{s,t} \right).
\end{align}

As with the spatial model, the spatial random effects ($\omega_s$) are assumed
to be drawn from a multivariate normal distribution with some covariance matrix
$\bm{\Sigma}_\omega$:

\begin{equation}
\bm{\omega} \sim \mathrm{MVNormal} \left( \bm{0}, \bm{\Sigma}_\omega \right),
\end{equation}

and we assume the same for the spatiotemporal random effects, with each time
slice being given its own independent set of random effects ($\bm{\epsilon}_t$)
with covariance matrix $\bm{\Sigma}_{\epsilon,t}$:

\begin{equation}
\bm{\epsilon}_t \sim \mathrm{MVNormal} \left( \bm{0}, \bm{\Sigma}_{\epsilon,t} \right).
\end{equation}

The spatial random effects account for spatial factors that are constant across
time, for example, depth and substrate type. The spatiotemporal random effects
account for factors that vary from year to year spatially such as bottom
temperature, water circulation patterns, species interactions, and species
movement.

Here we use Pacific Cod as an example species to illustrate the model
components. The approach includes the same generation of spatial 'knots' as in
the spatial model described in Appendix \@ref(app:spatial-modeling) (Figure
\@ref(fig:sdmTMB-spt-spde)). We use 200 knots, which for spatial coverage of
the synoptic trawl surveys, seems to be of adequately high resolution to
capture the spatial and spatiotemporal variation. We tested this assumption by
increasing the number of knots for a selection of species and ensuring that the
estimated trends did not qualitatively differ.

We can project predictions from the model to a fine-scale (2 km $\times$ 2 km)
grid using the covariance projection matrix (Figures
\@ref(fig:sdmTMB-spt-plot-all-effects-sqrt), \@ref(fig:sdmTMB-spt-plot-all-effects-sqrt-depth)). We can also look at the
individual components of the model. For the models without fixed effect
predictors for depth, the fixed effect predictions each year are constant
spatially (Figure \@ref(fig:sdmTMB-spt-plot-fixed-effects)). The spatial random
effects are constant across years (Figure
\@ref(fig:sdmTMB-spt-plot-spatial-effects)) and the spatiotemporal random effects
vary across years (Figure \@ref(fig:sdmTMB-spt-plot-spatiotemporal-effects)).
We can look at residuals through space and time to check if there appears to be
remaining spatial or spatiotemporal autocorrelation (Figure
\@ref(fig:sdmTMB-spt-spatial-residuals)).

We can then calculate expected biomass $B_t$ in year $t$ as:

\begin{equation}
B_t = \sum_{j = 1}^{n_j}
  w_j \cdot \exp \left( \bm{X}_{j,t} \bm{\beta} + \omega_j + \epsilon_{j,t} \right),
\end{equation}

where $j$ references a 2 km $\times$ 2 km grid cell within the survey domain
and $w_j$ represents the weight or area of that grid cell (4 km^2^) (Figure
\@ref(fig:sdmTMB-plot-index)). In other words, we sum the predicted biomass
across all grid cells within the survey domain for each year. We generated
standard errors on the annual estimates of log biomass via the generalized
Delta method as implemented in TMB [@kristensen2016]. Similar to the findings
of @thorson2015, we found that the model-based biomass index trends often had
lower CVs (coefficient of variations) than the design-based index trends and
often helped stabilize biomass estimates for outlying years compared to the
design-based index trends (e.g. Figure \@ref(fig:sdmTMB-plot-index)).

We found little difference in the predicted biomass index between models that
included or did not include depth covariates (e.g. Figure
\@ref(fig:sdmTMB-plot-index)). We found that the main difference between models
with or without depth covariates was in the models with depth covariates having
slightly more spatially resolved estimates of biomass (Figure
\@ref(fig:sdmTMB-spt-plot-all-effects-sqrt) vs. Figure
\@ref(fig:sdmTMB-spt-plot-all-effects-sqrt-depth)).
We chose to not include depth and depth squared as predictors in the main
biomass index illustrated throughout this report. Whereas in nearly all cases
including depth covariates generated a similar index to excluding them, in
a few cases including them generated what appeared to be unrealistic deviations
in biomass when a combination of the quadratic shape of the relationship
between depth and biomass combined to generate exceedingly high over low
estimates of biomass on the border of the survey polygons. This remains a topic
of research and will be investigated by the authors in the future.

```{r sdmTMB-spt-data}
library(dplyr)
library(ggplot2)
library(sdmTMB)

survey <- "SYN QCS"
n_knots <- 200
bias_correct <- FALSE
# dat <- readRDS(here::here("report/data-cache/walleye-pollock.rds"))
dat <- readRDS(here::here("report/data-cache/pacific-cod.rds"))
dat <- gfplot:::tidy_survey_sets(dat$survey_sets, survey, years = seq(1, 1e6),
  density_column = "density_kgpm2")
dat <- mutate(dat, density = density * 1000 * 1000)
dat <- gfplot:::interp_survey_bathymetry(dat)$data
dat <- gfplot:::scale_survey_predictors(dat)

grid_locs <- gfplot:::make_prediction_grid(
  filter(dat, year == max(dat$year)), survey = survey,
  cell_width = 2)$grid
grid_locs <- rename(grid_locs, depth = akima_depth)
grid_locs$year <- NULL
```

```{r sdmTMB-spt-spde, fig.cap='Example triangularization mesh for Pacific Cod and the Queen Charlotte Sound survey across all years. Red dots represent the knot locations and open black circles represent the location of the survey sets.', fig.width=9, out.width="3.7in"}
spde <- make_spde(dat$X, dat$Y, n_knots = n_knots)
plot_spde(spde)
```

```{r sdmTMB-spt-model}
m_st <- sdmTMB(
  formula = density ~ 0 + as.factor(year),
  data = dat, time = "year", spde = spde, family = tweedie(link = "log"),
  silent = TRUE)
predictions <- predict(m_st, newdata = grid_locs)
```

```{r sdmTMB-spt-model-depth}
m_st_depth <- sdmTMB(
  formula = density ~ 0 + as.factor(year) + depth_scaled + depth_scaled2,
  data = dat, time = "year", spde = spde, family = tweedie(link = "log"),
  silent = TRUE)
predictions_depth <- predict(m_st_depth, newdata = grid_locs)
```

```{r sdmTMB-spt-residuals, fig.cap='Randomized quintile residual histogram.'}
dat$resids <- residuals(m_st) # randomized quantile residuals
# gfplot::plot_qres_histogram(dat$resids)

dat$resids_depth <- residuals(m_st_depth) # randomized quantile residuals
# gfplot::plot_qres_histogram(dat$resids_depth)
```

```{r sdmTMB-spt-plot-function}
plot_map_spt <- function(dat, column) {
  ggplot(dat, aes_string("X", "Y", fill = column)) +
    geom_raster() +
    facet_wrap(~year) +
    coord_fixed() +
    labs(x = 'Easting', y = 'Northing')
}
```

```{r sdmTMB-spt-plot-all-effects-sqrt, fig.cap='Predictions from geostatistical spatiotemporal model for Pacific Cod in Queen Charlotte Sound. Predicted biomass is shown with a fourth-root-distributed colour scale.'}
ticks <- c(100, 500, 1000, 2000)
limits <- c(0, max(exp(predictions$data$est))*1.001)
plot_map_spt(predictions$data, "exp(est)") +
  ggtitle("Prediction (fixed effects + all random effects)") +
  scale_fill_viridis_c(trans = "fourth_root_power", option = "C",
    breaks = ticks, limits = limits) +
  labs(fill = 'Biomass estimate\nkg/km^2')
```

\clearpage

```{r sdmTMB-spt-plot-all-effects-sqrt-depth, fig.cap='Predictions from a geostatistical spatiotemporal model that includes the effects of depth and depth squared for Pacific Cod in Queen Charlotte Sound. Predicted biomass is shown with a fourth-root-distributed colour scale. Note the similarity to the previous figure which does not include the depth predictors.'}
plot_map_spt(predictions_depth$data, "exp(est)") +
  ggtitle("Prediction (fixed effects + all random effects)") +
  scale_fill_viridis_c(trans = "fourth_root_power", option = "C",
    breaks = ticks, limits = limits) +
  labs(fill = 'Biomass estimate\nkg/km^2')
```

\clearpage

```{r sdmTMB-spt-plot-fixed-effects, fig.cap='Fixed effect predictions from a geostatistical spatiotemporal model for Pacific Cod in Queen Charlotte Sound. Here the only fixed effects are the mean effects for each year resulting in fixed effect predictions that are same throughout the spatial region for each year.'}
plot_map_spt(predictions$data, "exp(est_fe)") +
  ggtitle("Prediction (fixed effects only)") +
  scale_fill_viridis_c(trans = "fourth_root_power", option = "C") +
  labs(fill = 'Mean estimate\nkg/km^2')
```

\clearpage

```{r sdmTMB-spt-plot-spatial-effects, fig.cap='Spatial random effect predictions from a geostatistical spatiotemporal model for Pacific Cod in Queen Charlotte Sound. The spatial random effects account for spatial factors that are constant across years, for example, depth and substrate type.'}
red_blue_fill <- scale_fill_gradient2(midpoint = 0,
    high = scales::muted("red"),
    mid = "white",
    low = scales::muted("blue"))

plot_map_spt(predictions$data, "est_re_s") +
  ggtitle("Spatial random effects only") +
  red_blue_fill +
  labs(fill = 'Deviation from\nexpected biomass\nin log space')
```

\clearpage

```{r sdmTMB-spt-plot-spatiotemporal-effects, fig.cap='Spatiotemporal random effect predictions from a geostatistical spatiotemporal model for Pacific Cod in Queen Charlotte Sound. The spatiotemporal random effects account for factors that vary from year to year spatially such as bottom temperature, water circulation patterns, and species interactions.'}
plot_map_spt(predictions$data, "est_re_st") +
  ggtitle("Spatiotemporal random effects only") +
  red_blue_fill +
  labs(fill = 'Deviation from\nexpected biomass\nin log space')
```

\clearpage

```{r sdmTMB-spt-spatial-residuals, fig.cap='Spatial residuals plotted through time on the link (log) scale.'}
red_blue_fill2 <- scale_colour_gradient2(midpoint = 0,
    high = scales::muted("red"),
    mid = "white",
    low = scales::muted("blue"))

ggplot(dat, aes(X, Y, col = resids)) +
  geom_point() + facet_wrap(~year) + coord_fixed() +
  labs(x = 'Easting', y = 'Northing', colour = "Residual") +
  red_blue_fill2
```

\clearpage

```{r sdmTMB-get-index}
 # not bias correcting for speed:
ind <- get_index(predictions, bias_correct = FALSE)
ind_depth <- get_index(predictions_depth, bias_correct = FALSE)
```

\clearpage

```{r sdmTMB-plot-index, fig.width=7, fig.cap='Relative biomass index predictions from the design-based approach, the geostatistical approach without covariates, and the geostatistical approach with depth covariates. Note the similarity between the two geostatistical models.'}
# ind_design <- readRDS(here::here('report/data-cache/walleye-pollock.rds'))
ind_design <- readRDS(here::here('report/data-cache/pacific-cod.rds'))
ind_design <- ind_design$survey_index
ind_design <- ind_design %>%
  filter(survey_abbrev  == 'SYN QCS') %>%
  mutate(
    est = biomass / exp(mean(log(biomass))),
    lwr = lowerci / exp(mean(log(biomass))),
    upr = upperci / exp(mean(log(biomass))),
    type = 'Design based'
  )

ind_geo <- ind %>%
  mutate(
    lwr = lwr / exp(mean(log(est))),
    upr = upr / exp(mean(log(est))),
    est = est / exp(mean(log(est))),
    type = "Geostatistical"
  )

ind_geo_depth <- ind_depth %>%
  mutate(
    lwr = lwr / exp(mean(log(est))),
    upr = upr / exp(mean(log(est))),
    est = est / exp(mean(log(est))),
    type = "Geostatistical (depth)"
  )

ind_combined <- bind_rows(ind_design, ind_geo) %>%
  bind_rows(ind_geo_depth)
ggplot(ind_combined, aes(year, est, fill = type)) +
  geom_line(aes(colour = type)) +
  geom_ribbon(aes(ymin = lwr, ymax = upr), alpha = 0.4) +
  xlab('Year') +
  ylab('Biomass estimate divided by geometric mean') +
  labs(fill = 'Type', colour = 'Type') +
  scale_fill_brewer(palette = "Dark2") +
  scale_colour_brewer(palette = "Dark2")
```

\clearpage
