# SURVEY BIOMASS INDEX TRENDS {#sec:survey-trend-models}

Fisheries and Oceans, Canada (DFO) conducts various fishery-independent
multispecies research surveys annually or biannually. These include four
synoptic bottom trawl surveys and two longline hook surveys. The survey areas
correspond roughly with Groundfish Management areas that fishery quotas are
allocated by British Columbia waters. For the synoptic trawl surveys, the West
Coast Vancouver Island survey corresponds to management areas 3CD, the Queen
Charlotte Strait survey corresponds to areas 5AB, the Hecate Strait surveys
correspond to areas 5CD, and the West Coast Haida Gwaii surveys corresponds to
area 5E. The Hard Bottom Longline survey is split into northern and southern
segments. The southern survey area corresponds to management areas 3CD and 5AB
while the northern survey area corresponds to areas 5CDE.

## SYNOPTIC BOTTOM TRAWL SURVEYS

DFO, together with the Canadian Groundfish Research and Conservation Society,
implemented a coordinated set of bottom trawl surveys that together cover the
continental shelf and upper slope of most of the BC coast. The surveys follow
a random depth stratified design and use the same bottom trawl fishing gear and
fishing protocols [@sinclair2003syn]. The surveys were designed to provide
a synopsis of all species available to bottom trawl gear as opposed to focusing
on specific species. There are a total of four synoptic (SYN) surveys: Hecate Strait
(HS), West Coast Vancouver Island (WCVI), Queen Charlotte Sound (QCS), and West
Coast Haida Gwaii (WCHG) (Figure \@ref(fig:intro-maps)). Surveys have been
conducted on both chartered commercial vessels and government research vessels.
The Hecate Strait survey and West Coast Vancouver Island survey have been
conducted on the Canadian Coast Guard research trawler WE Ricker while the
Queen Charlotte Sound survey has been conducted on chartered commercial fishing
vessels. Two of the synoptic surveys are conducted each year on an alternating
basis so that each survey is conducted every two years.

## HARD BOTTOM LONGLINE SURVEYS

The Pacific Halibut Management Association, in consultation with DFO, initiated
a depth-stratified, random design research longline survey conducted with
chartered commercial fishing vessels in 2006. The survey employs standardized
longline snap gear and fishing methods and alternates annually between the
northern and southern portions of BC. The survey is designed to provide catch
rates of all species and biological samples of rockfish from the outside
coastal waters of BC for stock assessment, alternating between northern and
southern areas.

## INTERNATIONAL PACIFIC HALIBUT COMMISSION FISHERY INDEPENDENT SURVEY

The International Pacific Halibut Commission’s (IPHC) fishery independent
survey is the longest times series of longline survey data in BC. This survey
is a fixed station survey that has been conducted annually, with chartered
commercial fishing vessels deploying fixed gear, in Canadian waters (IPHC Area
2B), since 1963. It provides distribution, biomass, age, growth and maturity
data that are used in the annual assessment of Pacific Halibut
(\emph{Hippoglossus stenolepis}). In 2003, the IPHC provided the opportunity to
deploy an additional technician to enumerate and identify catch to species on
a hook-by-hook basis and to collect biological data on rockfishes during their
2B survey operations (e.g. [@flemming2012iphc]). The complete enumeration of
species during the IPHC survey was recorded in 1995, 1996, and in all survey
years beginning in 2003. In the years between 1995 and 2003, and in 2013,
regular species composition sampling occurred over the first 20 hooks (20%) on
each survey skate of gear.

We note that the IPHC survey targets Halibut and the survey timing is aligned
with when Halibut are growing, which may increase hook competition and could
produce potential biases. Furthermore, when a set line is longer (made up of
more skates) it may fall outside of the targeted bottom type and therefore may
have a different catch composition than directly on the targeted station.

Details on the design of the various surveys referenced in this report can be
found in the following documents:

1. Synoptic Survey, Queen Charlotte Sound (SYN QCS): @williams2018synqcs

1. Synoptic Survey, West Coast Vancouver Island (SYN WCVI): @nottingham2017synwcvi

1. Synoptic Survey, Hecate Strait (SYN HS): @wyeth2018synhs

1. Synoptic Survey, West Coast Haida Gwaii (SYN WCHG): @williams2018synwchg

1. Hard Bottom Longline Survey, Outside (HBLL OUT): @dfo2018ifmp

1. Hard Bottom Longline Survey, Inside (HBLL INS): @lochead2007irf

1. Hecate Strait Multispecies Assemblage Survey (MSA HS): @choromanski2004hsmsa

1. International Pacific Halibut Commission fishery independent survey (IPHC FISS): @flemming2012iphc

The main biomass index trends illustrated on the figure pages represent the
design-based index trends that have historically been used in Pacific
Biological Station groundfish stock assessments. We calculated the trawl and
longline survey relative biomass index trends from the same design-based
bootstrap approach used in recent BC groundfish stock assessment reports. The
code to perform the calculations was originally written by Norm Olsen at
Pacific Biological Station and is automatically applied to the available survey
data to populate the GFBio database. We extracted the relative biomass values
directly from GFBio. Nonetheless, we have included the relevant equations below
for clarity. We also present geostatistical model based estimates of biomass
index trends for the trawl surveys for comparison Section
(\@ref(sec:geostatistical)).

## TRAWL SURVEY DESIGN-BASED ESTIMATION {#sec:trawl-methods}

For all trawl surveys, and for a given species, we calculated the relative
biomass density $B$ in year $y$ as:

\begin{equation}
  B_y = \sum_{i = 1}^k C_{y,i} A_i
\end{equation}

where $C_{y,i}$ represents the mean CPUE in kg/km^2^ for the species in year
$y$ and stratum $i$, $A_i$ represents the area of stratum $i$ in km^2^, and $k$
represents the number of strata. We calculated the CPUE ($C_{y,i}$) for a given
species in year $y$ and stratum $i$ as:

\begin{equation}
  C_{y,i} = \frac{\sum_{j = 1}^{n_{y,i}} \left( W_{y,j} / D_{y,j} w_{y,j}\right)}{n_{y,i}}
\end{equation}

where $W_{y,j}$ represents the catch weight (kg) for the species in year $y$,
stratum $i$, and tow $j$; $D_{y,j}$ represents the distance travelled in km by
tow $j$ in year $y$; $w_{y,j}$ represents the net opening width in km for year
$y$ and tow $j$; and $n_{y,i}$ represents the number of tows in year $y$ and
stratum $i$.

## LONG-LINE SURVEY DESIGN-BASED ESTIMATION {#sec:ll-methods}

For the HBLL surveys, and for a given species, we calculated the relative
biomass density $B$ in year $y$ as:

\begin{equation}
  B_y = \sum_{i = 1}^k C_{y,i} A_i
\end{equation}

where $C_{y,i}$ represents the mean CPUE in parts (fish) per
km\textsuperscript{2} for the species in year $y$ and stratum $i$, $A_i$
represents the area of stratum $i$ in km^2^, and $k$ represents
the number of strata. We calculated the CPUE ($C_{y,i}$) for a given species in
year $y$ and stratum $i$ as:

\begin{equation}
  C_{y,i} = \frac{\sum_{j = 1}^{n_{y,i}} \left( N_{y,j} / H_{y,j} w_{y,j}\right)}{n_{y,i}}
\end{equation}

where $N_{y,j}$ represents the number of fish caught for the species in year
$y$, stratum $i$, and set $j$; $H_{y,j}$ represents the number of hooks $\times$
the hook spacing in km in set $j$ in year $y$; $w_{y,j}$ represents an arbitrary
swept width of 30 feet or 0.009144 km for year $y$ and tow $j$; and $n_{y,i}$
represents the number of sets in year $y$ and stratum $i$. The hook spacing is
8 feet or 0.0024384 km for
the inside and outside HBLL surveys.

Details on the design-based estimation of biomass density index for the IPHC
survey are shown in Section (TODO).

## DESIGN-BASED-INDEX CONFIDENCE INTERVALS

We calculated bootstrap confidence intervals on $B_y$ by repeatedly calculating
$B_y$ given the above equations but each time re-sampling, with replacement,
from the available tows within each strata. We drew 1000 bootstrap replicates
of $B_y$, $B_y^{\mathrm{rep}}$, and calculated 95% quantile-based confidence
intervals on $B_y^{\mathrm{rep}}$.

## GEOSTATISTICAL SPATIOTEMPORAL BIOMASS INDEX TRENDS {#sec:geostatistical}

TODO: Bring these notes together into cohesive paragraphs.

In recent years there has been a movement toward model-based standardization
of survey biomass index trends (e.g. REFS). This approach has a theoretical
advantage over a design-based approach for a number of reasons ... TODO.

For comparison, we include model-based index trends for the synoptic trawl
surveys. We include them as a point of comparison so the reader can gauge when
the two approaches most differ. Large differences are likely a result of the
random positioning of the survey sets for a given year, which can lead to
occasional large production deviations and biomass from year to year or
large uncertainties on the biomass estimate for certain years.

Authors of groundfish stock assessments may want to consider model-based index
standardization on a case-by-case basis.

We use the geostatistical model as described in Section
(\@ref(sec:spatial-modeling)) with the addition of spatiotemporal random
effects ($\epsilon_{s,t}$) and annual predictors for the mean biomass each year:

\begin{align}
  y_{s,t} &\sim \mathrm{Tweedie}(\mu_{s,t}, p, \phi), \quad 1 < p < 2,\\
  \mu_{s,t} &= \exp \left( \bm{X}_{s,t} \bm{\beta} + \omega_s + \epsilon_{s,t} \right),
\end{align}

as with the spatial model, the spatial random effects ($\omega_s$) are assumed
drawn from a multivariate normal distribution with some covariance matrix
$\bm{\Sigma}_\omega$:

$$\bm{\omega} \sim \mathrm{MVNormal} \left( \bm{0}, \bm{\Sigma}_\omega \right),$$

and we assume the same for the spatiotemporal random effects, with each time
slice being given its own independent set of random effects:

$$\bm{\epsilon_t} \sim \mathrm{MVNormal} \left( \bm{0}, \bm{\Sigma}_{\epsilon,t} \right).$$

TODO: Describe equations

The spatial random effects account for spatial factors that are constant
through year, for example, depth and substrate type. The spatiotemporal random
effects account for factors that vary from year to year spatially such as
bottom temperature, circulation patterns, species interactions, and species
movement.

The approach includes the same generation of spatial 'knots' to make the
problem computationally feasible through the SPDE (stochastic partial
differential equation) approximation and a predictive process model (Figure
\@ref(fig:sdmTMB-spt-spde)). Here, we use 200 knots, which for spatial coverage
of the synoptic trawl surveys, seems to be of adequately high resolution to
capture the spatial and spatiotemporal variation. We tested this assumption by
increasing the number of knots for a selection of species and ensuring that the
estimated trends did not qualitatively differ.

We use Pacific Cod as an example species to illustrate the model components.

We can project predictions from the model to a fine-scale (2 $\times$ 2 km) grid
using the covariance projection matrix generated by INLA... (Figures
\@ref(fig:sdmTMB-spt-plot-all-effects-sqrt)).

We can look at residuals through space and time to check if there appears to be
remaining spatial or spatiotemporal autocorrelation (Figure
\@ref(fig:sdmTMB-spt-spatial-residuals)).

We chose to not include depth and depth squared as predictors in this model for
simplicity and because it made little difference to the biomass index
estimates. While in nearly all cases including depth covariates generated
a similar index, in a few cases it generated what appeared to be unrealistic
estimates of biomass when a combination of the quadratic shape of the
relationship between depth and biomass combined to generate exceedingly high
estimates of biomass on the border of the survey polygons. This remains a topic
of research and will be investigated by the authors in the future.

We can look at the individual components of the model. For the models without
fixed effect predictors for depth, the fixed effect predictions each year are
constant spatially (Figure \@ref(fig:sdmTMB-spt-plot-fixed-effects)). The
spatial random effects are constant across years (Figure
\@ref(fig:sdmTMB-spt-plot-fixed-effects)) and the spatiotemporal random effects
vary across years (Figure \@ref(fig:sdmTMB-spt-plot-spatiotemporal-effects)).
By summing the predicted biomass across all grid cells within the survey
boundary, we can generate a biomass estimate through time (Figure
\@ref(fig:sdmTMB-plot-index)). We generated the standard errors on the annual
estimates of log biomass via the generalized Delta method as implemented in
TMB.

Similar to the findings of @thorson2015, we find that the model-based
biomass index trends often have lower CVs than the design-based index trends
and often help stabilize biomass estimates for outlying years in the
design-based index trends (e.g. Figure \@ref(fig:sdmTMB-plot-index)).

We find little difference in the predicted biomass index for most stocks (e.g.
Figure \@ref(fig:sdmTMB-plot-index)). The main difference lies in the model
with depth predictors having slightly more spatially resolved estimates of
biomass in space (Figure \@ref(fig:sdmTMB-spt-plot-all-effects-sqrt) vs. Figure
\@ref(fig:sdmTMB-spt-plot-all-effects-sqrt-depth)).

\clearpage

<!-- We can look at the spatial random effects but represent consistent deviations in space through time that are not accounted for by our fixed effects of depth. In other words, these deviations represent consistent biotic and abiotic factors that are affecting biomass density but are not accounted for in the model. -->

<!-- And finally we can look at the spatiotemporal random effects that represent deviation from the fixed effect predictions and the spatial random effect deviations. These represent biotic and abiotic factors that are changing through time and are not accounted for in the model. -->

```{r sdmTMB-spt-data}
library(dplyr)
library(ggplot2)
library(sdmTMB)

survey <- "SYN QCS"
n_knots <- 200
bias_correct <- FALSE
# dat <- readRDS(here::here("report/data-cache/walleye-pollock.rds"))
dat <- readRDS(here::here("report/data-cache/pacific-cod.rds"))
dat <- gfplot:::tidy_survey_sets(dat$survey_sets, survey, years = seq(1, 1e6),
  density_column = "density_kgpm2")
dat <- mutate(dat, density = density * 1000 * 1000)
dat <- gfplot:::interp_survey_bathymetry(dat)$data
dat <- gfplot:::scale_survey_predictors(dat)

grid_locs <- gfplot:::make_prediction_grid(
  filter(dat, year == max(dat$year)), survey = survey,
  cell_width = 2)$grid
grid_locs <- rename(grid_locs, depth = akima_depth)
grid_locs$year <- NULL
```

```{r sdmTMB-spt-spde, fig.cap='Example triangularization mesh for Pacific Cod and the Queen Charlotte Sound survey across all years. Red dots represent the knot locations and open black circles represent the location of the survey sets.', fig.width=9, out.width="3.7in"}
spde <- make_spde(dat$X, dat$Y, n_knots = n_knots)
plot_spde(spde)
```

```{r sdmTMB-spt-model}
m_st <- sdmTMB(
  formula = density ~ 0 + as.factor(year),
  data = dat, time = "year", spde = spde, family = tweedie(link = "log"),
  silent = TRUE)
predictions <- predict(m_st, newdata = grid_locs)
```

```{r sdmTMB-spt-model-depth}
m_st_depth <- sdmTMB(
  formula = density ~ 0 + as.factor(year) + depth_scaled + depth_scaled2,
  data = dat, time = "year", spde = spde, family = tweedie(link = "log"),
  silent = TRUE)
predictions_depth <- predict(m_st_depth, newdata = grid_locs)
```

```{r sdmTMB-spt-residuals, fig.cap='Randomized quintile residual histogram.'}
dat$resids <- residuals(m_st) # randomized quantile residuals
# gfplot::plot_qres_histogram(dat$resids)

dat$resids_depth <- residuals(m_st_depth) # randomized quantile residuals
# gfplot::plot_qres_histogram(dat$resids_depth)
```

```{r sdmTMB-spt-plot-function}
plot_map_spt <- function(dat, column) {
  ggplot(dat, aes_string("X", "Y", fill = column)) +
    geom_raster() +
    facet_wrap(~year) +
    coord_fixed() +
    labs(x = 'Easting', y = 'Northing')
}
```

```{r sdmTMB-spt-plot-all-effects-sqrt, fig.cap='Predictions from geostatistical spatiotemporal model for Pacific Cod in Queen Charlotte Sound. Predicted biomass is shown with a fourth-root-distributed colour scale.'}
ticks <- c(100, 500, 1000, 2000)
limits <- c(0, max(exp(predictions$data$est))*1.001)
plot_map_spt(predictions$data, "exp(est)") +
  ggtitle("Prediction (fixed effects + all random effects)") +
  scale_fill_viridis_c(trans = "fourth_root_power", option = "C",
    breaks = ticks, limits = limits) +
  labs(fill = 'Biomass estimate\nkg/km^2')
```

\clearpage

```{r sdmTMB-spt-plot-all-effects-sqrt-depth, fig.cap='Predictions from a geostatistical spatiotemporal model that includes the effects of depth and depth squared for Pacific Cod in Queen Charlotte Sound. Predicted biomass is shown with a fourth-root-distributed colour scale. Note the similarity to the previous figure which does not include the depth predictors.'}
plot_map_spt(predictions_depth$data, "exp(est)") +
  ggtitle("Prediction (fixed effects + all random effects)") +
  scale_fill_viridis_c(trans = "fourth_root_power", option = "C",
    breaks = ticks, limits = limits) +
  labs(fill = 'Biomass estimate\nkg/km^2')
```

\clearpage

```{r sdmTMB-spt-plot-fixed-effects, fig.cap='Fixed effect predictions from a geostatistical spatiotemporal model for Pacific Cod in Queen Charlotte Sound. Here the only fixed effects are the mean effects for each year resulting in fixed effect predictions that are same throughout the spatial region for each year.'}
plot_map_spt(predictions$data, "exp(est_fe)") +
  ggtitle("Prediction (fixed effects only)") +
  scale_fill_viridis_c(trans = "fourth_root_power", option = "C") +
  labs(fill = 'Mean estimate\nkg/km^2')
```

\clearpage

```{r sdmTMB-spt-plot-spatial-effects, fig.cap='Spatial random effect predictions from a geostatistical spatiotemporal model for Pacific Cod in Queen Charlotte Sound. The spatial random effects account for spatial factors that are constant across years, for example, depth and substrate type.'}
red_blue_fill <- scale_fill_gradient2(midpoint = 0,
    high = scales::muted("red"),
    mid = "white",
    low = scales::muted("blue"))

plot_map_spt(predictions$data, "est_re_s") +
  ggtitle("Spatial random effects only") +
  red_blue_fill +
  labs(fill = 'Deviation from\nexpected biomass\nin log space')
```

\clearpage

```{r sdmTMB-spt-plot-spatiotemporal-effects, fig.cap='Spatiotemporal random effect predictions from a geostatistical spatiotemporal model for Pacific Cod in Queen Charlotte Sound. The spatiotemporal random effects account for factors that vary from year to year spatially such as bottom temperature, water circulation patterns, and species interactions.'}
plot_map_spt(predictions$data, "est_re_st") +
  ggtitle("Spatiotemporal random effects only") +
  red_blue_fill +
  labs(fill = 'Deviation from\nexpected biomass\nin log space')
```

\clearpage

```{r sdmTMB-spt-spatial-residuals, fig.cap='Spatial residuals plotted through time on the link (log) scale.'}
red_blue_fill2 <- scale_colour_gradient2(midpoint = 0,
    high = scales::muted("red"),
    mid = "white",
    low = scales::muted("blue"))

ggplot(dat, aes(X, Y, col = resids)) +
  geom_point() + facet_wrap(~year) + coord_fixed() +
  labs(x = 'Easting', y = 'Northing', colour = "Residual") +
  red_blue_fill2
```

\clearpage

```{r sdmTMB-get-index}
 # not bias correcting for speed:
ind <- get_index(predictions, bias_correct = FALSE)
ind_depth <- get_index(predictions_depth, bias_correct = FALSE)
```

\clearpage

```{r sdmTMB-plot-index, fig.width=7, fig.cap='Relative biomass index predictions from the design based approach, the geostatistical approach without covariates, and the geostatistical approach with depth covariates. Note the similarity between the two geostatistical models.'}
# ind_design <- readRDS(here::here('report/data-cache/walleye-pollock.rds'))
ind_design <- readRDS(here::here('report/data-cache/pacific-cod.rds'))
ind_design <- ind_design$survey_index
ind_design <- ind_design %>%
  filter(survey_abbrev  == 'SYN QCS') %>%
  mutate(
    est = biomass / exp(mean(log(biomass))),
    lwr = lowerci / exp(mean(log(biomass))),
    upr = upperci / exp(mean(log(biomass))),
    type = 'Design based'
  )

ind_geo <- ind %>%
  mutate(
    lwr = lwr / exp(mean(log(est))),
    upr = upr / exp(mean(log(est))),
    est = est / exp(mean(log(est))),
    type = "Geostatistical"
  )

ind_geo_depth <- ind_depth %>%
  mutate(
    lwr = lwr / exp(mean(log(est))),
    upr = upr / exp(mean(log(est))),
    est = est / exp(mean(log(est))),
    type = "Geostatistical (depth)"
  )

ind_combined <- bind_rows(ind_design, ind_geo) %>%
  bind_rows(ind_geo_depth)
ggplot(ind_combined, aes(year, est, fill = type)) +
  geom_line(aes(colour = type)) +
  geom_ribbon(aes(ymin = lwr, ymax = upr), alpha = 0.4) +
  xlab('Year') +
  ylab('Biomass estimate divided by geometric mean') +
  labs(fill = 'Type', colour = 'Type') +
  scale_fill_brewer(palette = "Dark2") +
  scale_colour_brewer(palette = "Dark2")
```

\clearpage
