\clearpage

\vspace{0mm}

\clearpage

# IPHC SURVEY INDEX {#app:iphc-survey-index}

```{r iphc-examples}
yyr_set_counts <- readRDS(paste0(file.path(dc, "iphc/yelloweye-rockfish.rds")))$set_counts
dog_set_counts <- readRDS(paste0(file.path(dc, "iphc/north-pacific-spiny-dogfish.rds")))$set_counts

bait_set_counts <- readRDS(paste0(file.path(dc, "iphc/hook-with-bait.rds")))$set_counts

```


The International Pacific Halibut Commission (IPHC) conducts an annual stock
assessment longline survey in waters from California to Alaska, including
British Columbia waters [@flemming2012iphc]. The survey's main goal is to
provide data on Pacific Halibut (*Hippoglosus stenolepis*) for stock
assessment purposes. 

At each station, the fishing gear consists of a set of skates each of about 100
hooks. Up to eight skates are on each set, with the number of skates per set
varying between years. For each set the IPHC calculates an 'effective skate
number', which we use here to scale the count of each species of interest
and obtain a catch rate
for each set (described below). The effective skate number "standardizes survey
data in years when the number of hooks, hook spacing, or hook type varied"
[@yamanaka2008iphc]. An effective skate of one represents a skate of 100 circle
hooks with 18-foot spacing [@yamanaka2008iphc].

For British Columbia waters, the survey has enumerated non-halibut species
since 1995 (to varying degrees of species identification). 
For each species, the catch rate of a set is the number of individuals caught 
per effective skate.
We bootstrap these catch rates within each year to
to give annual bootstrapped means, bias-corrected and adjusted
(BCa) bootstrapped 95\% confidence intervals, and bootstrapped coefficients of
variation (CV).

However, complications arise because of differing data
collection protocols in different years. 
We would like a survey index that spans as long a time period as possible, and,
ideally, also covers all the coastwide waters off British Columbia (excluding
the Strait of Georgia which the IPHC survey does not enter). Although the
spatial coverage and the technical details of the survey are not consistent from
year to year (as described below), we attempt to construct a survey index for
as many species as possible for as long a time period as possible. We also 
determine whether each index can be considered representative of all British 
Columbia waters. For each species, the resulting index gives what we term
the International Pacific Halibut Commission fishery independent survey series
(IPHC FISS).

The approach taken is described below, and builds on that developed for
assessments
of Redbanded Rockfish [@edwards2017redbanded] and Yelloweye Rockfish
[@yamanaka2018yelloweyeoutside]. The Redbanded assessment was the first to
develop an abundance index from the IPHC survey that went back to 1995, and
included data up to 2012. For the Yelloweye assessment the methods
were extended to demonstrate that the index based on waters north of Vancouver
Island could be considered representative of the coastwide population. 

TODO: Put in main text: This work demonstrates that a consistent index can be constructed despite the
survey design changing through the years. 

## IPHC DATA 

In British Columbia waters (IPHC area 2B), since 2003 a third observer has been
deployed on the IPHC survey to identify all catch to the species level on a
hook-by-hook basis and to conduct biological sampling [@flemming2012iphc], although
in 2013 there was no such observer. Observers were also deployed prior to 2003,
although data are not available in such detail, as summarised in
Table \@ref(tab:IPHCdata). For some years only the first 20&nbsp;hooks from each skate
were enumerated, and for other years all hooks were enumerated but the data are
only available at the set level (i.e.&nbsp;we do not know which hook caught which
species, only how many individuals from each species were caught on the whole
set). The data were extracted from various spreadsheets and the DFO database
`GFBio`, and all originally came from the IPHC. For only some of the years
were the locations off the west coast of Vancouver Island (WCVI) sampled. Note
that, for simplicity we use the term 'first 20&nbsp;hooks' since samplers on the
vessels generally targetted the first 20&nbsp;hooks deployed from each skate.
However, for
operational reasons (particularly in areas of high catch rates), sometimes the
20&nbsp;hooks would come from elsewhere within a skate, but would always consist of
20 consecutive hooks (e.g. @dykstra2002iphc, @dykstra2002iphc).
TODO - force new paragraph.

<!-- 
Table:  (\#tab:IPHCdata) Summary of available data from the IPHC stock assessment
longline surveys. 'Data resolution' indicates at what level the data are
available, and 'WCVI?' indicates whether or not the survey included locations
off the west coast of Vancouver Island. 'Location of data' indicates where the
data were accessed from, either our DFO GFBio database or from spreadsheets.
^1^For 1995, the biological data were in the file
`1995_IPHC_SSA_Rockfish_catch_from_Kelly_Ames.xls` on DFO's Inshore
Rockfish shared drive, and effective skates were obtained from Aaron Ranta
(IPHC) in the file `1995EffSktValues by Station.xlsx`. ^2^For 1996-2002, the
data were in the file `2B AllSpecies 96-02 roundIII.xls`, which originally
came from the IPHC. ^3^For 2013 the data were in the file `2013 20-Hook
Data.xls`, which originally came from the IPHC. For easier access, the data
from spreadsheets are now all included in our gfplot R package.

Year        Hooks enumerated         Data resolution   Location of data      WCVI?
----------- ------------------------ ----------------- -------------------- -------
1995        All                      Set-by-set        Spreadsheets^1^        N
1996        All                      Set-by-set        Spreadsheet^2^         N
1997-1998   First 20 of each skate   Set-by-set        Spreadsheet^2^         N
1999        First 20 of each skate   Set-by-set        Spreadsheet^2^         Y
2000        First 20 of each skate   Set-by-set        Spreadsheet^2^         N
2001-2002   First 20 of each skate   Set-by-set        Spreadsheet^2^         Y
2003-2011   All                      Hook-by-hook      DFO database `GFBio`   Y
2012        All (bait experiment)    Hook-by-hook      DFO database `GFBio`   Y
2013        First 20 of each skate   Set-by-set        Spreadsheet^3^         Y
2014-2017   All                      Hook-by-hook      DFO database `GFBio`   Y
-->

\begin{table}[t]
\centering
\caption{Summary of available data from the IPHC stock assessment longline
surveys. 'Data resolution' indicates at what level the data are available, and
'WCVI?' indicates whether or not the survey included locations off the west
coast of Vancouver Island. 'Location of data' indicates where the data were
accessed from, either our DFO {\tt GFBio} database or from spreadsheets. $^1$For
1995, the biological data were in the file
"1995\_IPHC\_SSA\_Rockfish\_catch\_from\_Kelly\_Ames.xls" on DFO's Inshore
Rockfish shared drive, and effective skates were obtained from Aaron Ranta
(IPHC) in the file "1995EffSktValues by Station.xlsx". $^2$For 1996-2002, the
data were in the file "2B AllSpecies 96-02 roundIII.xls", which originally
came from the IPHC. $^3$For 2013 the data were in the file "2013 20-Hook
Data.xls", which originally came from the IPHC. For easier access, the data from
spreadsheets are now all included in our gfplot package.}
\label{tab:IPHCdata}
\begin{tabular}{llllc}
  \hline
Year & Hooks enumerated & Data resolution & Location of data & WCVI?\\
  \hline
1995 & All & Set-by-set & Spreadsheets$^1$ & N\\
1996 & All & Set-by-set & Spreadsheet$^2$ & N\\
1997-1998 & First 20 of each skate & Set-by-set & Spreadsheet$^2$ & N\\
1999 & First 20 of each skate & Set-by-set & Spreadsheet$^2$ & Y\\
2000 & First 20 of each skate & Set-by-set & Spreadsheet$^2$ & N\\
2001-2002 & First 20 of each skate & Set-by-set & Spreadsheet$^2$ & Y\\
2003-2011 & All & Hook-by-hook & DFO database {\tt GFBio} & Y\\
2012 & All (bait experiment)      & Hook-by-hook & DFO database {\tt GFBio} & Y\\
2013 & First 20 of each skate & Set-by-set & Spreadsheet$^3$ & Y\\
2014-2017 & All & Hook-by-hook & DFO database {\tt GFBio} & Y\\
   \hline
\end{tabular}
\end{table}


From Table \@ref(tab:IPHCdata), four issues are apparent: 

1. For 1997-2002 and 2013 only the first 20&nbsp;hooks of each skate were
enumerated, whereas for all other years all hooks were enumerated. Thus, the
data from each year cannot simply be considered as comparable and analysed as
one consecutive time series.

1. For the datasets for 1995, 1996, 1997-2002 and 2013, data are only
available at the set-by-set level, in terms of numbers of a given species per
effective skate. Which species was caught on each hook is not available, unlike
for 2003-2012 and 2014-2107. Thus, for 1995 and 1996 we cannot calculate catch rates
based on the first 20&nbsp;hooks (because we only have set-by-set level data),
whereas we can do that for 2003-2012 and 2014-2017, and the 20-hook data is the only
information we have for 1997-2002 and 2013.

1. In 2012 a bait experiment was conducted such that data from all skates
could not be used; see Section \@ref(sec:chum).

1. The WCVI was not visited in every year, so the spatial coverage is not
consistent across years.

\begin{table}[t] \centering
\caption{Summary of how the four Series {\bf A}, {\bf B}, {\bf C} and {\bf D}
are constructed. Numbers in parentheses indicate the number of years for which
data for each Series are available. 'Only north of WCVI' indicates Series that
only consider stations north of Vancouver Island (thus excluding those off the
WCVI), 'Full coast' indicates Series that use all stations from the whole
coast. The rows indicate how many hooks the catch rates for each Series are
based on.}
\label{tab:seriesSumm}
\begin{tabular}{rcc}
\hline
 & Only north of WCVI & Full coast \\
\hline
First 20 hooks from each skate & {\bf A}(21) & {\bf D}(18)\\
All hooks from each skate & {\bf B}(16) & {\bf C}(14)\\
\hline
~\\    % to add some whitespace before text carries on
\end{tabular}
\end{table}

To address issues 1, 2 and 4 we therefore construct four time series (whose
structure is summarised in Table \@ref(tab:seriesSumm)) for each species:
  
**Series A** -- 1997-2017 stations north of WCVI, with catch rates based on
first 20&nbsp;hooks only (which is all we have for 1997-2002 and 2013).
 
**Series B** -- 1995, 1996, 2003-2012 and 2014-2017 stations north of WCVI, with
catch rates based on all hooks (which is all we have for 1995 and 1996).

**Series C** -- 2003-2012 and 2014-2017 stations coastwide (including WCVI), with
catch rates based on all hooks.

**Series D** -- 1999 and 2001-2017 stations coastwide (including WCVI), with
catch rates based on first 20&nbsp;hooks only (which is all we have for 1999,
2001-2002 and 2013).

We would like to obtain an index series with as long a timespan as possible,
and, ideally, over as broad a geographic region as possible. Since Series&nbsp;A is
the longest time series, we take this and expand it to Series&nbsp;AB, defined as:

**Series AB** -- for stations north of WCVI, combine the 1995 and 1996 values
from Series B, based on all hooks, with the 1997-2017 values from Series A that
are based on first 20 hooks only. See Section&nbsp;\@ref(sec:iphc-equations).

The resulting Series&nbsp;AB covers the stations north of WCVI. In
Sections&nbsp;\@ref(sec:iphc-AD) and&nbsp;\@ref(sec:iphc-BC) we show how we determine, for
each species, 
whether we can consider
this series to be representative of the full coast (i.e.&nbsp;including the WCVI), by
comparing the series that exclude stations off the WCVI (Series&nbsp;A and&nbsp;B) with
those that include the stations off the WCVI (Series&nbsp;C and&nbsp;D, respectively).

## SPATIAL LOCATIONS OF STATIONS

For the IPHC survey, from 1995-1997 the stations were
arranged in Y-shapes; they were not exactly the same locations each year, but
fairly close to each other. Since 1998 
the stations have
been positioned equidistant from one another on a fixed 10-nautical-mile square grid
[@flemming2012iphc]. In 1999 the survey first
went to the WCVI, did not go there in 2000, but has since 2001 onwards. See 
@edwards2017redbanded and 
@yamanaka2018yelloweyeoutside for maps that demonstrate the different coverage
(and that show stations that caught Yelloweye and Redbanded Rockfish, respectively).

Given the difference in coverage between years, for Series&nbsp;A and&nbsp;B we exclude
those stations south of 50.6$^{\circ}$ latitude, which is near the northern tip
of Vancouver Island. This
latitude was chosen so that all the stations from 1995-1997 are included. The
stations for 1995-1997 show good overlap (north of Vancouver Island) with the
stations from 1998 onwards, despite not being on the exact same 10-nautical-mile
square grid [@yamanaka2018yelloweyeoutside].
Series&nbsp;C and&nbsp;D use all stations coastwide (by definition).

Each year, a few stations may be declared unusable by the IPHC and are excluded 
from our analyses (e.g. the 
hook-tally sheet got blown overboard for station 2113 in the year 2008), with
usability codes described in Table\ \@ref(tab:iphc-set-codes). In particular,
we include stations deemed as 'Usable but omit from any geospatial analysis',
but these should be exluded for complex spatial analyses.

```{r iphc-hook-codes, tab.cap="Description of hook observation classifications and corresponding codes in theGFBio database.", eval=FALSE}

hookCodes = cbind("Description"=c("Unknown", "Empty hook",
    "Bait on hook",
    "Animal on hook (fish or invertebrate)", "Species head on hook",
    "Species droped off hook", "Bait skin on hook", "Hook not observed",
    "Eaten or bitten (by shark, etc.)"),
    "HOOK_YIELD_CODE" = 0:8)      # from iphc0314.Snw
csasdown::csas_table(hookCodes)
```

\begin{table}[t] \centering
\caption{Description of hook observation classifications and corresponding codes
in theGFBio database.}
\label{tab:seriesSumm}
\begin{tabular}{lc}
\hline
Description & HOOK\_YIELD\_CODE\\
\hline
Unknown & 0\\
Empty hook & 1\\
Bait on hook & 2\\
Animal on hook (fish or invertebrate) & 3\\
Species head on hook & 4\\
Species droped off hook & 5\\
Bait skin on hook & 6\\
Hook not observed & 7\\
Eaten or bitten (by shark, etc.) & 8\\
\hline
\end{tabular}
\end{table}



\begin{table}[tbp]
\centering
\caption{Description of classification of IPHC sets, and indication
of which we include in our analyses.}
\label{tab:iphc-set-codes}
\begin{tabular}{lcc}
\hline
Description & U.USABILITY\_CODE & Included here? \\
\hline
Fully usable & 1 & Y\\
Usable but omit from swept area calculations & 21 & N\\
Usable but removed due to re-definition of survey area & 22 & N\\
Usable but omit from any biomass calculations & 27 & N\\
Usable but omit from any geospatial analysis & 52 & Y\\
\hline
\end{tabular}
\end{table}



## CHUM SALMON BAIT EXPERIMENT {#sec:chum}

Prior to 2012, Chum Salmon (\emph{Oncorhynchus keta}) was used for bait. But in
2012, a bait experiment was conducted [@henry2013iphc]. At each station three
different bait types were used on the same set: a consecutive four-skate Chum
Salmon treatment, a one-skate Pink Salmon (*Oncorhynchus gorbuscha*)
treatment, and a one-skate Walleye Pollock (*Theragra chalcogramma*)
treatment. The location of the three treatments on each set was randomized
throughout the survey, and each treatment was separated by one skate (1,800&nbsp;ft)
of hookless groundline. For consistency with previous years, we only consider
the four skates that used Chum Salmon as bait.

The effective skate number provided by the IPHC is for all skates used, which in
2012 will include skates that were not baited with Chum Salmon (Eric Soderlund,
IPHC, Seattle, WA, USA, pers.&nbsp;comm.). But we wish to only include the Chum
Salmon baited skates, and so we need to modify the effective skate number (see
below). The effective skate number depends on the number of observed hooks (Eric
Soderlund, IPHC, Seattle, WA, USA, pers.&nbsp;comm.), rather than the number of hooks
that were deployed. The bait experiment has not been repeated.




## CATCH RATE EQUATIONS {#sec:iphc-equations}

### CATCH RATE BASED ON ALL CHUM-BAIT HOOKS

For each species of interest, we wish to obtain a catch rate index which, 
for each year, will be the mean catch rate across all sets that year.
The units will be numbers of individuals
caught per effective skate. We only want to consider hooks that used
Chum Salmon as bait (hereafter 'chum-bait hooks'), because we have no
information as to how catch rates change depending on the bait
used. For our data, 2012 was the only year that hooks were not exclusively
chum-bait hooks.

Define:

$H_{it}$ -- number of observed chum-bait hooks in set $i$ in year $t$,

$H_{it}^*$ -- number of observed hooks for all bait types
($H_{it} \neq H_{it}^*$ only for 2012),

$E_{it}$ -- effective skate number of set $i$ in year $t$, which needs to be
based on observed chum-bait hooks,

$E_{it}'$ -- effective skate number from IPHC, which is based on all observed
hooks (regardless of bait).

Thus, $E_{it}$ is
\begin{equation}
E_{it} = \dfrac{H_{it}}{H_{it}^*} E_{it}'.
\end{equation}

Adapting equations on page 3 of [@yamanaka2008iphc], define:

$N_{it}$ -- the number of fish of a given species caught on set $i=1,2,...,n_t$
in year $t$, based on observed chum-bait hooks,

$n_t$ -- the number of sets in year $t$,

$C_{it}$ -- catch rate (with units of numbers per effective skate) for set $i$
in year $t$, based on observed chum-bait hooks, given by
\begin{equation}
C_{it} = \frac{N_{it}}{E_{it}}.
\label{catchPerSet}
\end{equation}
The catch rate index for year $t$, $I_t$ (numbers per effective skate), is then
the mean catch rate across all sets:
\begin{equation}
I_{t} = \frac{1}{n_t} \sum_{i=1}^{n_t} C_{it} = \frac{1}{n_t} \sum_{i=1}^{n_t} \frac{N_{it}}{E_{it}}.
\label{index}
\end{equation}

### CATCH RATE BASED THE FIRST 20 CHUM-BAIT HOOKS OF EACH SKATE

Let $\tilde{X}$ indicate a calculation of some value $X$ that is based only on
the first 20 hooks of each skate. These are the first 20 *numbered* hooks,
not the first 20 *observed* hooks (so not all of the numbered hooks may
have been observed). Thus we have:

$\tilde{H}_{it}$ -- number of observed chum-bait hooks in the first 20 hooks of
all skates in set&nbsp;$i$ in year&nbsp;$t$, 

$\tilde{E}_{it}$ -- effective skate number of set&nbsp;$i$ in year&nbsp;$t$ based on the
first 20 chum-bait hooks that were sent out on each skate.

Since effective skate number is a linear function of the number of hooks in a
set [@yamanaka2008iphc], we have
\begin{equation}
\tilde{E}_{it} = \dfrac{\tilde{H}_{it}}{H_{it}} E_{it} 
  \left( = \dfrac{\tilde{H}_{it}}{H_{it}^*} E_{it}' \right).
\label{effSkateScale}
\end{equation}

The resulting notation for the index will be:

$\tilde{I}_{t}$ -- catch rate index for year $t$ (in numbers of individuals per
effective skate) based on only the first 20 hooks sent out for each skate,

$\tilde{N}_{it}$ -- the number of individuals caught on set $i=1,2,...,n_t$ in year
$t$, based on observed chum-bait hooks and only the first 20 hooks sent out for
each skate,

$\tilde{C}_{it}$ -- catch rate (with units of numbers per effective skate) for
set $i$ in year $t$, based only on the first 20 hooks of each skate (and only
skates with chum as bait), such that 
\begin{equation}
\tilde{C}_{it} =
  \frac{\tilde{N}_{it}}{\tilde{E}_{it}}.
(\#eq:catchPerSet20)
\end{equation}
The catch rate index for year $t$,
$\tilde{I}_t$ (in units of numbers per effective skate), based on only
the first 20 hooks of each skate, is then the mean catch rate across all sets:
\begin{equation}
\tilde{I}_t = \frac{1}{\tilde{n}_t} 
  \sum_{i=1}^{\tilde{n}_t} \tilde{C}_{it} 
  = \frac{1}{\tilde{n}_t} \sum_{i=1}^{\tilde{n}_t} \frac{\tilde{N}_{it}}{\tilde{E}_{it}}.
\label{index20}
\end{equation}
  
We base calculations on bootstrapped means, and so $I_t$ and $\tilde{I}_t$ are 
calculate, for each year, by resampling the catch rates ($C_{it}$ 
or $\tilde{C}_{it}$) 10,000 times, and calculating a bootstrapped mean and
95\% bias-corected and adjusted confidence interval. 

### EQUIVALENCY OF CATCH RATES BASED ON ALL HOOKS AND ON JUST THE FIRST 20 HOOKS

Equation \@ref(eq:catchPerSet20) can be written as 
\begin{equation}
\tilde{C}_{it} = \frac{\tilde{N}_{it}}{\tilde{E}_{it}} = 
  \frac{H_{it}}{\tilde{H}_{it}} \frac{\tilde{N}_{it}}{E_{it}}.
\label{tildeCatchPerSet}
\end{equation}
If all hooks are equally likely to catch an individual of the given species,
then the catch rates based on the first 20 hooks of each skate should be an 
unbiased sample of the catch rates based on all the hooks.
The ratio of individuals caught, $\tilde{N}_{it} / N_{it}$, should equal (on average)
the ratio of hook numbers, $\tilde{H}_{it} / H_{it}$, because a proportionally
reduced number of fish are caught on the proportionally fewer hooks. Thus
\begin{equation}
\dfrac{\tilde{H}_{it}}{H_{it}} = \dfrac{\tilde{N}_{it}}{N_{it}}
\end{equation}
such that
\begin{equation}
\tilde{C}_{it} = \dfrac{N_{it}}{\tilde{N}_{it}} \frac{\tilde{N}_{it}}{E_{it}} = \frac{N_{it}}{E_{it}} = C_{it}.
\end{equation}

If the catch rates are greatly different, then this suggests that the catch
rates from the first 20&nbsp;hooks are not equivalent to the catch rates based on all
the hooks. This is why we compare Series&nbsp;A and Series&nbsp;B.  

### CONSTRUCTING SERIES AB

For each species, we wish to join up the 1995 and 1996 data from Series&nbsp;B (based on all
hooks) to the 1997-2017 data from Series&nbsp;A. 
The 1995 and 1996 data are only available as numbers
of individuals caught for all hooks, and not as numbers caught in the first 20
hooks. For 1997-2002 and 2013 we only have numbers caught for the
first 20 hooks. But for 2003-2012 and 2014 onwards we have hook-by-hook data,
and so can
compute catch rates for all hooks or based on just the first 20 hooks
(i.e.&nbsp;these overlapping years are the only years that contribute to both
Series&nbsp;A and Series&nbsp;B).

<!-- 
Figure caption from YYR:  \twofig{bcaKeepSerAB1}{bcaKeepSerAB2}{(a) Each of the two catch rate series from Figure \ref{fig:bcaKeepSerA} is divided by the geometric mean of its bootstrapped annual means (with the geometric mean based on the overlapping years only). (b) The catch rate index Series&nbsp;AB that could be used as a model input (although the index in Appendix&nbsp;C was used in this stock assessment). Series&nbsp;AB extends the original Series&nbsp;A by incorporating the suitably scaled 1995 and 1996 values from Series&nbsp;B (see text).}
-->

For Series&nbsp;A, define $G_A$ to be the geometric mean of the bootstrapped annual
means, with the geometric mean based only on the overlapping years (2003-2012
and 2014 onwards). Define $G_B$ similarly for Series B. By dividing the bootstrapped
values for each series by their respective geometric means, we obtain Figure


TODO for TODOexample. see the example in Figure&nbsp;TODO(a) for dogfish??. 
This shows that the rescaled Series A and Series B
are very similar for the overlapping years. Thus, on this scale, the 1995 and
1996 values from Series B look comparable to the full Series A data.

We statistically test the comparability by conducting a 
paired t-test [@crawley2002] on the scaled annual means for the overlapping years, to test the null
hypothesis that there is no difference between the scaled annual means, with 
resulting $p$-value defined as $p_{AB}$.
If $p_{AB} \geq 0.05$ then we cannot reject the null hypothesis. Then we join
up the two series in FigureTODO(a) by taking the rescaled 1995 and 1996 values
from Series&nbsp;B and joining to the rescaled Series&nbsp;A. Equivalently,
we just multiply the original Series-B means and confidence intervals for 1995 
and 1996 by 
$G_A/G_B$, and then join them up to the original Series&nbsp;A, as in Figure TODO(b), to get
the longest time series possible, based on the first 20&nbsp;hooks from each
skate, rescaled for 1995 and 1996 for which numbers for the first 20&nbsp;hooks
are not available.

If $p_{AB} < 0.05$ then we cannot join the series, and so we stick with
Series&nbsp;A as being the longest. 

TODO equivalent of YYR Figure B.13 for an example species (dogfish?).

For worked examples see @edwards2017redbanded and @yamanaka2018yelloweyeoutside, or the output for any
species (TODO give gfplot command if possible).

TODO Too much to go through a species, but show summary information table for
each year, number of usable sets coastwide and north of VI, 
mean effective skate number based on all hooks and based on
first 20, and mean hook competition adjustment factor if I calculate it.


### CONSTRUCTING SERIES&nbsp;D (20 HOOKS, COASTWIDE) AND COMPARING IT WITH SERIES&nbsp;A (20 HOOKS, NORTH OF VANCOUVER ISLAND) {#sec:iphc-AD}

We now consider Series&nbsp;D, which is for the first 20&nbsp;hooks of each
skate (like for Series&nbsp;A) but covers the whole coast, including the WCVI
(unlike Series&nbsp;A), as was summarised in Table\ \ref{tab:seriesSumm}. 

In the same way that we just compared Series&nbsp;A and Series&nbsp;B, we
divide each series by its geometric mean ($G_A$ or $G_D$, based on the
overlapping years of A and&nbsp;D) and conduct
a paired t-test on the scaled annual bootstrapped means for the overlapping
years
to give a $p$-value ($p_{AD}$).
Again, if $p_{AD} \geq 0.05$ then we consider the two series comparable.

For Series&nbsp;A and Series&nbsp;D this means that we can consider
the relative changes in Series&nbsp;A (that excludes WCVI) to be the same as
those in Series&nbsp;D (that includes the full coast), and hence we can consider
Series&nbsp;A to be representative of the full coast. So the population off the 
WCVI is not showing a different relative trend to the rest of the coast.

We do not need to join up the two series, we just wish to verify whether the
relative changes in Series&nbsp;A can be considered representative of the full
coast. If $p_{AD}<0.05$ then this is not the case.

For the last Yelloweye Rockfish assessment [@yamanaka2018yelloweyeoutside],
the relative scaled patterns for Series&nbsp;A and Series&nbsp;D appeared similar for the
overlapping years (though no statistical test was done).
But the *absolute* catch rates were different, 
with inclusion of the WCVI stations in Series&nbsp;D consistently reducing
the mean annual catch rates from those of Series&nbsp;A
(that did not include the WCVI stations),
with $G_A/G_D=1.12$ for the
overlapping years (it would equal 1 if the catch rates were the same).
The stations off the WCVI had lower average catch rates of Yelloweye Rockfish
than the remaining stations. 

So while inclusion of the
WCVI stations does not appear to change the \emph{relative} pattern of the index
of the population, it does change the absolute values. Therefore the stations
off the WCVI have to be included or excluded consistently to construct an index
series; since we have more years that do not have stations off the WCVI
(Table&nbsp;\ref{tab:seriesSumm}), we consistently exclude these stations (giving
Series&nbsp;A).

### CONSTRUCTING SERIES&nbsp;C (ALL HOOKS, COASTWIDE) AND COMPARING IT WITH SERIES&nbsp;B (ALL HOOKS, NORTH OF VANCOUVER ISLAND) {#sec:iphc-BC}

Similarly, we also construct Series&nbsp;C, which is for all hooks from each 
skate (like for
Series&nbsp;B) but covers the whole coast, including the WCVI (unlike Series&nbsp;B), as
was summarised in Table&nbsp;\ref{tab:seriesSumm}. 

We again compare each series scaled by its geometric mean of the overlapping
years and conduct
a paired t-test on the scaled annual bootstrapped means for the overlapping
years,
giving a $p$-value $p_{BC}$.
If $p_{BC}\geq 0.05$ then we consider the two series comparable.
This means that we can consider the relative changes in Series&nbsp;B to reflect
those of the full coast (Series&nbsp;C).

For the last Yelloweye Rockfish assessment [@yamanaka2018yelloweyeoutside],
similarly to Series&nbsp;A and&nbsp;D just discussed, the relative scaled patterns for Series&nbsp;B and Series&nbsp;C appeared similar for the
overlapping years.
But the *absolute* catch rates were different, 
with inclusion of the WCVI stations in Series&nbsp;C consistently reducing
the mean annual catch rates from those of Series&nbsp;B
(that did not include the WCVI stations),
with $G_B/G_C=1.11$ for the
overlapping years.

So the stations off the WCVI had lower average catch rates of Yelloweye Rockfish
than the remaining stations. This holds whether we look at all hooks (here) 
or just the first 20 by comparing Series&nbsp;A and&nbsp;D (above). 

### RESULTING LONGEST INDEX FROM THE IPHC SURVEY

So, if $p_{AB} \geq 0.05$ then Series&nbsp;AB can be created and that is the 
longest IPHC survey index that can be constructed. If $p_{AD} \geq 0.05$ and
$p_{BC} \geq 0.05$ then Series&nbsp;A and Series&nbsp;B can be considered
representatitive of the full coast, and thus so can Series&nbsp;AB.

If if $p_{AB} \geq 0.05$ but either 
$p_{AD} < 0.05$ or
$p_{BC} < 0.05$
then Series&nbsp;AB is still the longest time series possible, but can
only be considered to represent the waters north of Vancouver Island, not
the full coast.

For some special cases of species that are rarely caught (or that weren't
specifically enumerated in 1995 and 1996, for example), then Series&nbsp;A may
be the longest, or even possibly Series&nbsp;C or&nbsp;D if the species was not
caught in some years. Some rare special cases may not yet be fully accounted for
in our code, and may need to be verified on a species-by-species basis,
particularly
to ascertain whether a given species was being actively identified in a given
year (resulting in a zero catch rate rather than 'no data').

## HOOK COMPETITION

TODO Hook competition occurs ......


[@clark2008iphc]

[@webster2011iphc]

TODO The Yelloweye
assessment also considered a model of hook competition TODO **mention Marie's if
I don't get anything in. Note, however, that this work was not directly
incorporated into the analysis in Appendix~C that was used as the survey index
in the assessment model.

Extending the earlier notation, define:

$N_{its}^{(0)}$ -- the number of fish of species $s$ that we would expect to
catch *in the absence of hook competition* on set $i=1,2,...,n_t$
in year $t$, based on observed chum-bait hooks,

$D_{its}$ -- true local density of species $s$ around set $i$ in year $t$,

$F_{its}$ -- the local rate of capture of bait by species $s$ around set $i$ in
year $t$.

The number $N_{its}^{(0)}$ is proportional to the true local density, $D_{its}$, and so
is what we wish to use as an index of abundance. And, by definition,
\begin{equation}
N_{its}^{(0)} = F_{its} H_{it},
\label{eq:Nits0}
\end{equation}
where, as earlier, $H_{it}$ is the number of observed chum-bait hooks in set $i$
in year $t$.


Based on Clark:

"Mathematically the process of baits being removed from a longline by different
species is the same as the process of fish being removed from a population by
different fisheries and natural predators". Each species removes a certain
proportion of the baits per unit time, so the Baranov catch equation can be used
to give
\begin{equation}
N_{its} = F_{its} H_{it} \frac{\left(1 - \mathrm{e}^{- Z_{it}} \right)}{Z_{it}},
\label{eq:Nits}
\end{equation}
where $Z_{it}$ is the sum of the instantaneous rates of capture by all species, i.e.
\begin{equation}
Z_{it} = \sum_s F_{its},
\label{eq:Zit}
\end{equation}
and the soak time can be left out because there is no significant difference
between shorter and longer soak times (the sets soak for at least five hours).

Substituting $F_{its} H_{it}$ into \@ref(eq:Nits) from \@ref(eq:Nits0) gives
\begin{equation}
N_{its} = \frac{N_{its}^{(0)} \left(1 - \mathrm{e}^{- Z_{it}} \right)}{Z_{it}},
\end{equation}
which upon rearranging gives
\begin{equation}
N_{its}^{(0)} = \frac{N_{its} Z_{it}}{1 - \mathrm{e}^{- Z_{it}}},
\label{eq:Nits0rearrange}
\end{equation}

Define $P_{it}$ to be the proportion of observed chum-bait hooks
(for set $i$ in year $t$) that still have the bait on them (and are therefore
assumed to be continuously actively fishing). The remaining baits are either 
captured by a fish or lost TODO plus other options from table  ,  and considering
lost bait to be another 'species', 
\begin{equation}
P_{it} = 1 - \frac{\sum_s N_{its}}{H_{it}}
\end{equation}
such that 
\begin{equation}
\sum_s N_{its} = {H_{it}} \left(1 -  P_{it} \right).
\label{eq:sumsNits}
\end{equation}


Now, summing \@ref(eq:Nits) over all species $s$ gives
\begin{equation}
\sum_s N_{its} = \sum_s F_{its} H_{it} \frac{\left(1 - \mathrm{e}^{- Z_{it}} \right)}{Z_{it}}
\end{equation}
Substituting from \@ref(eq:sumsNits) and \@ref(eq:Zit) gives
\begin{align}
{{H_{it}}} \left(1 -  P_{it} \right) & = Z_{it} H_{it}
  \frac{\left(1 - \mathrm{e}^{- Z_{it}} \right)}{Z_{it}} \\
1 -  P_{it} & = 
  1 - \mathrm{e}^{- Z_{it}} \\
Z_{it} & = - \ln P_{it}.
\end{align}

....Sub 19 and 20 into 14, to give the competition adjustment factor...... do in morning....

\begin{align}
N_{its}^{(0)} & = \frac{-\ln P_{it}}{1 - P_{it}} N_{its} \\
              & = A_{it} N_{its},
\label{eq:Nits0rearrange}
\end{align}
where 
\begin{equation}
A_{it} = \frac{-\ln P_{it}}{1 - P_{it}},
\end{equation}
is the competition adjustment factor for each set in each year. It
scales up the observed number of each species caught to give the expected number
of species caught (accounting for hook competition), depending on the proportion $P_{it}$ of observed hooks
that are returned still with bait on them.

Note that, by definition, $0 \leq P_{it} \leq 1$, and so $\ln P_{it} \leq 0$. As
$P_{it} \rightarrow 1$ then by L'H\mbox{\^o}pital's Rule 
<!--may need a REF, Kreyszig doesn't have it-->
\begin{align}
\lim_{P_{it} \rightarrow 1} A_{it} & = \lim_{P_{it} \rightarrow 1} \frac{-\ln P_{it}}{1 - P_{it}}\\
 & = \dfrac{\lim_{P_{it} \rightarrow 1} (-1 / P_{it})}{\lim_{P_{it} \rightarrow 1} (-1)}\\
  & = 1,
\end{align}
such that $N_{its}^{(0)} = N_{its}$, which equals $0$ (because if all 
all hooks are returned with bait on then $N_{its} = 0$ 
for all species $s$).  

\ 


However, as $P_{it} \rightarrow 0$, $A_{it} \rightarrow
\infty$, such that the expected number $N_{its}^{(0)} \rightarrow \infty$.
In practice this
happens fairly often. There are 
`r nrow(bait_set_counts)`
sets in the data. Of the 
`r nrow(filter(bait_set_counts, !is.na(N_it)))` 
for which all hooks on each skate were enumerated,
`r nrow(filter(bait_set_counts, N_it == 0))`
had $P_{it} = 0$. Considering the 
`r nrow(filter(bait_set_counts, !is.na(N_it20)))`
sets for which we can calculate catch rates based on the first 20\ hooks of each
skate (i.e.\ all years except 1995 and 1996),
`r nrow(filter(bait_set_counts, N_it20 == 0))`
had $P_{it}=0$ for the first 20\ hooks.

Therefore the choice of what to use for $A_{it}$ when $P_{it} = 0$ is important.
One option would be to set it to the value obtained if only one hook with bait
is returned for a set. For a set with 800 observed hooks (essentially the
maximum), the smallest possible positive value is $P_{it} = 1/800$, which gives
$A_{it} = 6.69$.

```{r iphc-hook-comp-adjustment, fig.cap="Plot of how the hook competition adjustment factor varies with the proportion of hooks that are returned with bait, with the lowest proportion set to 1/800."}
Pit <- c( seq(1/800, 0.1, length=200), seq(0.1  , 0.99, 0.01) )
hook_adjust = - log(Pit) / (1 - Pit)
plot(Pit, hook_adjust, ylim=c(0, max(hook_adjust)), type="l",
     xlab = expression("Proportion of hooks with bait, P"[it]),
     ylab = "Hook competition adjustment factor")
````

```{r iphc-hook-with-bait, warning=FALSE, width=5, height=6, fig.cap="Histograms of number of hooks returned with baits in each set, for each year in which all hooks were enumerated."}
ggplot(data=bait_set_counts,
       aes(N_it) ) +
    xlab("Number of hooks returned each year") +
    ylab("Frequency") +    
    geom_histogram() +
    facet_wrap(. ~ year, ncol=4)
```

```{r iphc-hook-with-bait-twenty, warning=FALSE, width=5, fig.cap="Histograms of number of hooks out of the first 20 of each skate that were returned with bait, for each year (except 1995 and 1996 for which this cannot be calculated)."}
ggplot(data=bait_set_counts,
       aes(N_it20)) +
    xlab("Number of hooks returned each year") +
    ylab("Frequency") +    
    geom_histogram() +
    facet_wrap(. ~ year, ncol=4)
# be good to have the 20 hook ones in different colour, or just do all N_it20 except for 1995 and 1996
```


check that, see notes on p6, plot if time, code into gfplot..... save as separate folder

No: And $F_{its}$ is also
proportional to $D_{its}$, such that






The



\clearpage
