# COMPUTATIONAL ENVIRONMENT {-}

This version of the document was generated on `r Sys.time()` with
`r devtools::session_info()$platform$version` [@r2018] and R package versions:

```{r env, echo = FALSE, results='asis', message=FALSE, warning=FALSE, eval=TRUE, cache=FALSE}
library(TMB)
library(rstan)
library(bookdown)
library(rmarkdown)
pkgs <- sort(c("broom", "gfplot", "ggplot2", "dplyr", "rstan", "knitr", "glmmTMB",
  "TMB", "broom.mixed", "scales", "purrr", "xtable", "kableExtra", "cowplot", "csasdown", "bookdown", "rmarkdown"))
info <- devtools::session_info()$packages
info %>% dplyr::filter(package %in% pkgs) %>%
  dplyr::select(package, version, date, source) %>%
  dplyr::rename(Package = package, Version = version, Date = date, Source = source) %>%
  csasdown::csas_table() %>%
  kable_styling(latex_options = c("hold_position"))
```

In particular, most of the data extraction and many of the plots were made with
the gfplot package version 0.1.1. The document was compiled with R package
csasdown version 0.0.1. The specific versions used to generate this report can
be installed with:

```{r, eval=FALSE, echo=TRUE}
devtools::install_github("pbs-assess/gfplot", ref = "f55710d")
devtools::install_github("pbs-assess/csasdown", ref = "baf78d8")
```

\clearpage
