# RELATIONSHIPS BETWEEN FISH SURVEY DENSITY AND DEPTH {#app:depth-curves}

```{r sdmTMB-depth-all-calculations}
surveys <- c("SYN QCS", "SYN HS", "SYN WCHG", "SYN WCVI")
fi <- list.files(here::here("report/map-cache/synoptic"), full.names = FALSE)
out <- purrr::map_df(fi, function(i) {
  mm <- readRDS(here::here(paste0("report/map-cache/synoptic/", i)))
  out <- purrr::map_df(1:4, function(ii) {
    if (length(mm$models[[ii]]$models) > 1L) {
      rd <- dplyr::filter(mm$raw_dat, survey == surveys[ii])
      if (!'depth_mean' %in% names(rd))
        stop('Scaling mean and SD are missing.')
      range_d <- -1 * exp(rd$depth_mean[1] +
          range(rd$depth_scaled, na.rm = TRUE) * rd$depth_sd[1])
      x <- seq(-3, 4, length.out = 300)
      x2 <- x^2
      B <- mm$models[[ii]]$models$model$par
      if (B[[3]] <= 0) { # quadratic must be :-( shaped
        y <- B[[1]] + x * B[[2]] + x2 * B[[3]]
      } else {
        y <- NA
      }
      out <- data.frame(
        depth = -1*exp(x * rd$depth_sd[1] + rd$depth_mean[1]),
        y = exp(y) * 1000, # convert to kg/km^2 (was scaled by 1000 already)
        survey = surveys[ii],
        species = gsub("-", " ", gsub(".rds", "", i)),
        stringsAsFactors = FALSE)
      out <- mutate(out,
        extrapolated = depth < min(range_d) * 1 |
          depth > max(range_d) * 1)
      out
    }
  })
  out
})

surveys <- c('HBLL OUT N', 'HBLL OUT S')
fi <- list.files(here::here("report/map-cache/hbll"), full.names = FALSE)
out_hbll <- purrr::map_df(fi, function(i) {
  mm <- readRDS(here::here(paste0("report/map-cache/hbll/", i)))
  out <- purrr::map_df(1:length(surveys), function(ii) {
    if (length(mm$models[[ii]]$models) > 1L) {
      rd <- dplyr::filter(mm$raw_dat, survey == surveys[ii])
      if (!'depth_mean' %in% names(rd))
        stop('Scaling mean and SD are missing.')
      range_d <- -1 * exp(rd$depth_mean[1] +
          range(rd$depth_scaled, na.rm = TRUE) * rd$depth_sd[1])
      x <- seq(-3, 4, length.out = 300)
      x2 <- x^2
      B <- mm$models[[ii]]$models$model$par
      if (B[[3]] <= 0) { # quadratic must be :-( shaped
        y <- B[[1]] + x * B[[2]] + x2 * B[[3]]
      } else {
        y <- NA
      }
      y <- B[[1]] + x * B[[2]] + x2 * B[[3]]
      out <- data.frame(
        depth = -1*exp(x * rd$depth_sd[1] + rd$depth_mean[1]),
        y = exp(y) / 100, # convert to 100 fish/km^2
        survey = surveys[ii],
        species = gsub("-", " ", gsub(".rds", "", i)),
        stringsAsFactors = FALSE)
      out <- mutate(out,
        extrapolated = depth < min(range_d) * 1 |
          depth > max(range_d) * 1)
      out
    }
  })
  out
})
out <- bind_rows(out, out_hbll)
dd <- out %>% group_by(species, survey) %>%
  # mutate(y = y / max(y))
  mutate(max_y = max(y[!extrapolated])) %>%
  mutate(y = ifelse(y < max_y * 1.15, y, NA)) %>%
  mutate(mode_depth = depth[y == max(y)[1]]) %>%
  group_by(species) %>%
  mutate(mean_mode_depth = mean(mode_depth)) %>%
  filter(depth >= -800)

if(french){
  dd$species <- en2fr(gfsynopsis:::first_cap(dd$species), french)
}

make_depth_plot <- function(.data, 
  ylab = if(french){
    expression(Densité~de~la~biomasse~du~relevé~(kg/km^2))
  }
  else{
    expression(Survey~biomass~density~(kg/km^2))
  }
  ,
  xlim = c(-800, 0)) {
  ggplot(.data, aes_string('depth', 'y', colour = 'survey')) +
    geom_line(lty = 2) +
    coord_cartesian(xlim = xlim) +
    labs(x = paste(en2fr('Depth', french), ' (m)'),
      y = ylab, colour = en2fr('Survey', french)) +
    # facet_wrap(~forcats::fct_reorder(species,
    #   mean_mode_depth),
    facet_wrap(~gfsynopsis:::first_cap(species), scales = "free_y", ncol = 4) +
    geom_line(data = dplyr::filter(.data, !extrapolated), lwd = 0.9) +
    scale_color_brewer(palette = "Dark2")
}
```

(ref:depth-relationship1) Predicted relationships between depth and biomass
density for all species across the four synoptic surveys (Part 1 of 2). Solid
lines indicate relationships predicted within the depth range of the survey and
dashed lines indicate extrapolated relationships beyond the observed depth.
These relationships are derived from the depth coefficients in the spatial
models that generate the map plots (e.g., Figure \@ref(fig:survey-maps)). These
plots provide a visual indication of which surveys encompass the entirety of the
depth distribution for particular species. See Appendix
\@ref(app:spatial-modeling) for details on the models underlying these
predictions.

```{r sdmTMB-depth-all-plots1, fig.width=10, fig.asp=1.10, out.width="\\textwidth", fig.cap="(ref:depth-relationship1)", fig.pos="H"}
.sp <- filter(dd, grepl('SYN', survey)) %>% 
  pull(species) %>% unique()
synoptic_data <- filter(dd, grepl('SYN', survey)) %>% 
  mutate(survey = as.factor(as.character(survey)))
filter(synoptic_data, species %in% .sp[1:36]) %>%
  make_depth_plot()
```

\clearpage

(ref:depth-relationship2) Part 2 of Figure \@ref(fig:sdmTMB-depth-all-plots1). Caption is the same otherwise.

```{r sdmTMB-depth-all-plots2, fig.width=10, fig.asp=1.3, out.width="\\textwidth", fig.cap="(ref:depth-relationship2)", fig.pos="p"}
filter(synoptic_data, species %in% .sp[37:length(.sp)]) %>%
  make_depth_plot()
```

\clearpage

(ref:depth-relationship-hbll) Same as Figure \@ref(fig:sdmTMB-depth-all-plots1) but for the outside hard bottom long line surveys. Caption is the same otherwise. Note that these panels do not extend as deep as the depths shown in Figures \@ref(fig:sdmTMB-depth-all-plots1) and \@ref(fig:sdmTMB-depth-all-plots2).

```{r sdmTMB-depth-all-plots1-hbll, fig.width=10, fig.asp=1.10, out.width="\\textwidth", fig.cap="(ref:depth-relationship-hbll)", fig.pos="H"}
.sp <- filter(dd, grepl('HBLL', survey)) %>% 
  pull(species) %>% unique()
hbll_data <- filter(dd, grepl('HBLL', survey)) %>% 
  mutate(survey = as.factor(as.character(survey)))
filter(hbll_data, species %in% .sp) %>%
  make_depth_plot(
    ylab = if(french){
    expression(Densité~du~relevé~(100~poissons/km^2))
  }
  else{
    expression(Survey~density~(100~fish/km^2))
  }
  ,
    xlim = c(-250, 0)
  )
```
