---
title: "Title here"
author:
- "First author here"
date: |
  ^1^Pacific Biological Station, Fisheries and Oceans Canada
     Nanaimo, BC, V6T 6N7, Canada; sean.anderson@dfo-mpo.gc.ca\
output:
  bookdown::pdf_document2:
    number_sections: no
    toc: no
    template: "article.tex"
bibliography: "refs.bib"
csl: "cfjas.csl"
link-citations: yes
linkcolor: "blue"
urlcolor: "blue"
abstract: >
  The abstract will go here.
---

\clearpage

```{r, echo = FALSE, message = FALSE}
library(knitr)
library(here)
opts_chunk$set(echo = FALSE, message = FALSE, warnings = FALSE)
```

# Introduction

Ecological resource management needs the ability to monitor a lot of things on a frequent basis but this takes a lot of time, people, and money.
- more frequently monitored populations do better (would be wonderful if we could say this; reference?)
 (or less frequently monitored populations do worse - may be easier to find a ref for this)
- legislation such as ... covers stocks of economic, conservation, cultural, ... value
 (or certain proportions/all species monitored/sustainably fished/conserved... = demands on personnel/money)
- but even "less important" stocks can form an important part of ecosystems
- major ecological surprises happen more often than we expect
- there is an increasing need to frequently monitor (more) stocks and across a large swath of dimensions given the rapid environmental change and increased variability in general reduced predictability associated with climate change

Thankfully, the advent of modern sensor technologies, relational databases, automated reporting (perhaps this one comes later), drones, etc. has resulted in vast quantities of data that can aid with this monitoring, but the data often remains hidden either because of permission or privacy issues or just through obscurity (or lack of mechanism/capacity to summarise/report).
- give some examples of monitoring programs from across Canada and around the world
- give an example of privacy issues (fishery locations and privacy act)
- give an example of volumes of data generating obscurity

In British Columbia, Canada our groundfish fisheries face these very problems. We have a lot of stocks, and not a lot of time, people, or money. We have 100% observer or video monitoring of our fleet. And although the Canadian government is moving towards an open data direction, we aren't entirely there yet, some data can never be shared due to privacy laws, and even when it is shared it is not easily digested due to the vast quantities of it. Plus, it requires some domain specific knowledge to interpret.

As a solution to this, here we introduce a data synopsis report for 100 groundfish species in British Columbia, Canada. Our goals with this report are to:
- facilitate discussions related to stock prioritization, fieldwork planning, ...
- increase transparency about our data holdings
- ensure we and others are regularly reviewing the available data especially for stocks that are between assessments or have never been assessed 
- produce standardized tools, derived data sets, and visualizations that are useful elsewhere in, for example, assessments

# Methods

We chose to focus our report on species that are of commerical, recreational, conservation or First Nations interest, or are regularly caught in DFO surveys. Of the various fisheries independent surveys that DFO conducts, we show data from those that provide the greatest spatial and taxonomic coverage of the species in the report. We chose to focus on aspects of the data that are most important to stock assessment or indicators of stock dynamics, and display both spatial and temporal aspects of the data.

In designing the report we made the following design considerations:
- limit two pages per species so that all data for one species can be viewed at once
- show empty plots to emphasize missing data and maintain the same layout for all species to allow rapid interpretation after becoming familiar with the layout
- highlight uncertainty wherever possible
- assemble and display metadata alongside visualization
- group similar focused visualizations together
- maintain consistent colour meaning throughout and use colourblind proof colours where possible
- describe the common visualizations in great detail at the beginning so that the report pages are almost entirely visualization


We wanted the entire report to be fully automated and reproducible so that it could be easily updated in the future. We also wanted the report production to be fully transparent and traceable regarding how the data are extracted from DFO databases and treated prior to plotting. To accomplish this, we developed a family of R packages that facilitate going directly from our relational databases, to extracting necessary data, to tidying the necessary data, to fitting statistical models, to visualizing the data and statistical models, and to compiling the visualisations into a report format. The code for the report was written in a modular way so that individual functions could be useful for other purposes, for example in stock assessments requiring similar statistical models or to extract data for a diverse array of projects in a consistent, standardized way.

We settled on visualization components that covered trends in time and space for fisheries independent surveys, trends in time and space for fisheries catch and catch per unit effort. Visualizations of length, age, growth and maturity from fisheries independent surveys and commercial data sources. We additionally show a summary of all commercial and research biological sample data availabilty for each species by counts of specimens.

# Results

The resulting visualizations for an example stock (Figure 1 and 2).

Some title for the left page: (~temporal and spatial population and fishery trends?)
- relative biomass trends from fisheries independent surveys
- maps of relative biomass from most recent fisheries independent surveys
- trends in fisheries catches through time by gear type and region
- trends in trawl fisheries catch per unit effort through time by region (issues with longline data still need to be resolved)
- maps of fisheries catch painted effort by gear type for a recent window of time

Some title for the right page (something about samples; ~biological trends, growth & maturity, and data availability?):
- length frequencies by survey and year
- age frequencies by survey and year
- length age growth models
- weight length growth models
- age and length at maturity models
- maturity frequencies by stage and month
- counts of the number of available fish lengths, weights, maturities, ages, and aging structures, from all surveys and all commercial samples (to see what might be available)

# Discussion

Quick recap of what we did and why it is important

This was only possible because of the carefully organized and maintained relational databases
- emphasize the value of putting resources into developing and maintaining proper databases 
- emphasize the importance of data accessibility even just within an agency
- give examples of what happens when databases are well organized and open (RAM Legacy, GPDD, ...)


Caveats to what we did
- briefly summarize the ones in the report into one paragraph

Comparison to even tangentially related efforts
- mention data dashboards and their advantages and disadvantages
- reference Chris Rooper's Alaska work add Dan's work on the east coast (is Dan's work what Christie Whelan had referenced when this all started? Or is there something else on the east coast?)

Where else this would be useful
- nearly anywhere with lots of populations or processes to monitor and available data maintained in an organized fashion
- especially useful for agencies responsible for a lot of stocks without the capacity to assess as frequently as management or ecological theory would suggest is ideal
- could also be useful for monitoring environmental data, not just biological data, and not just fisheries

Conclusions and where we are going with this next
- this was a useful exercise for us; already become an indispensable reference for many people and the tools have helped conduct other assessments rapidly
- plan to update on an annual or biennial basis and examine larger structural changes on a less frequent basis
- plan to include more climate related visualization such as centre of gravity and mean depth through time (ha! where to fit this...?)
- ...

# Acknowledgements

Steering committee, technical working group, data folks, reviewers, maintainers of the R packages that made it possible, those that have collected the data, those that have funded the data collection, ...

\clearpage

# Figures

```{r page1, fig.cap = "Figure caption goes here."}
# include_graphics(here("figs/decades-and-booms-viridis.pdf"))
```

```{r page1, fig.cap = "Figure caption goes here."}
# include_graphics(here("figs/decades-and-booms-viridis.pdf"))
```

\clearpage

# References
