---
title: "Title here"
author: |
  Sean C. Anderson^1^, and
  Elise A. Keppel^1^
date: 
address: |
  ^1^Pacific Biological Station, Fisheries and Oceans Canada
     Nanaimo, BC, V6T 6N7, Canada; sean.anderson@dfo-mpo.gc.ca\
output:
  bookdown::pdf_document2:
    number_sections: no
    toc: no
    template: "article.tex"
bibliography: "refs.bib"
csl: "cfjas.csl"
link-citations: yes
linkcolor: "blue"
urlcolor: "blue"
abstract: >
  The abstract will go here.
---

\clearpage

```{r, echo = FALSE, message = FALSE}
library(knitr)
library(here)
opts_chunk$set(echo = FALSE, message = FALSE, warnings = FALSE)
```

# Introduction

Monitoring is a critical component of effective ecological resource management. Monitoring can be used to choose management actions or learn about systems through adaptive management [REFs; @possingham2012]. It can also be used to engage stakeholders or raise awareness about the state of ecological systems [e.g. @possingham2012]. Focussed monitoring programs, such as those targeting a specific populations or ecosystems, can help parameterize complex statistical models and provide deep insights into the mechanisms driving population or ecosystem status [e.g. REF]. At the same time, broad monitoring programs are essential to detecting ecological surprises and the "unknown unknowns" [e.g. @hilborn1987; @wintle2010] --- a purpose only becoming more important in an era of climate change.

Although modern technologies have resulted in the collection of vast quantities of data, turning that data into effective monitoring remains a challenge. Agencies that collect data often lack the resources to report on all facets of it in a timely fashion. Thankfully, the movement towards open data has made many raw data sets available, but processing it requires domain-specific knowledge (e.g., about sampling protocols), advanced skillsets, and time --- all factors that limit the ability for stakeholders (e.g., resource users or concerned citizens) to derive their own insights.

In British Columbia (BC), Canada, Fisheries and Oceans Canada (DFO) faces these very problems in monitoring its groundfish populations and fisheries. Government groundfish scientists hold the potential responsibility of providing science advice on over 100 species. The commercial groundfish fleet is subject to 100% on-board and dockside monitoring and numerous fisheries-independent trawl and longline surveys are conducted annually producing large quantities of potentially valuable data. At the same time, assessment scientists face severe limitations in personnel and time. As a result, most stocks are not regularly examined through stock assessment leaving the majority of monitoring data unreported. And, although the Canadian government is engaged in an open-data initiative, some of DFO's data holdings will never be shared due to privacy laws, and the data that are shared are not easily digested without extensive knowledge of the survey programs and advanced data processing skill sets.

Automated visualization is a key solution to these problems. Rather than just being an endpoint to data analysis, there is an increasing realization that visualization should be an "integral part of the scientific process" [@fox2011].

- visualization can communicate to the masses
- visualization can surprise
- visualization can facilitate rapid understanding
- visualization can engage and raise awareness
- visualization can help find errors in data sets
- something about data dashboards

<!-- - "Second, new visualization approaches and technologies are needed to reduce the time and costs associated with generating visualizations of increasingly large and complex data relationships [@fox2011]." [@michener2012] "Unfortunately, visualization too often becomes an end product of scientific analysis, rather than an exploration tool that scientists can use throughout the research life cycle. However, new database technologies, coupled with emerging Web-based technologies, may hold the key to lowering the cost of visualization generation and allow it to become a more integral part of the scientific process." -[@fox2011] *Changing the Equation on Scientific Data Visualization!* -->

Here, as part of a solution to our above-mentioned challenges, we introduce an automated data visualization report we term our "groundfish data synopsis report" for 113 groundfish species in British Columbia, Canada. Our goals with this report are manifold and include: 
(1) ensuring DFO scientists, managers, and all interested parties are regularly reviewing indices of population status for all stocks, 
(2) facilitating discussions related to stock prioritization and survey programs, 
(3) increasing transparency about our data holdings, and
(4) producing standardized tools, derived data sets, and visualizations that are useful elsewhere in, for example, assessments.

# Methods

We focussed our report on species of commerical, recreational, conservation, or First Nations interest, and those regularly caught in DFO surveys. We included both commercial and survey population trends and biological data in the report to give an overall snapshot of available information for each species.

Of the various fisheries independent surveys that DFO conducts, we show data from those that provide the greatest spatial and taxonomic coverage of the species in the report. We chose to focus on aspects of the data that are most important to stock assessment or indicators of stock dynamics, and display both spatial and temporal aspects of the data.

In designing the report we made the following design considerations:

1. Fit all plots on two pages per species so that all data for one species could be viewed at once.
2. Show empty plots to emphasize missing data and maintain the same layout for all species.
3. Group similar focused visualizations together.
4. Assemble and display metadata alongside visualization
5. Maintain consistent colour meaning throughout (e.g. data from a given survey always uses the same colour) and use colourblind-proof colours where possible.
6. Describe the visualizations in detail at the beginning of the report so that the species pages are almost entirely visual.
7. Highlight uncertainty wherever possible (e.g. through shading to indicate periods of increased uncertainty in reported catches).
8. Focus on data types relevant to the most species.

We wanted the entire report to be fully automated and reproducible so that it could be easily updated in the future. We also wanted the report production to be fully transparent and traceable regarding how the data are extracted from DFO databases and treated prior to plotting. To accomplish this, we developed a family of R packages that facilitate direct extraction of necessary data from our relational databases, tidying and arranging the necessary data for modelling or plotting, fitting statistical models, plotting the data and statistical models, and compiling the visualisations into a report format. The code for the report was written in a modular way so that individual functions could be useful for other purposes, for example in stock assessments requiring similar statistical models or to extract data for a diverse array of projects in a consistent, standardized way.

We settled on visualization components that covered trends in time and space for fisheries independent surveys and commercial fisheries catch and catch per unit effort; visualizations of temporal trends in length and age frequencies; and visualizations of biological parameters inlcuding length at age, weight at length, age at maturity, length at maturity and maturity frequencies from fisheries independent surveys and commercial data sources. We additionally show a summary of all commercial and research biological sample data availabilty for each species by counts of specimens.

Throughout report development we solicited feedback from stakeholders and other potential users of the report on data to include as well as on report structure and design. Feedback was also sought from data experts for various species groups to ensure that the data being captured were accurate, complete, and relevant.

# Results

The report includes two pages of visualization for 113 stocks of British Columbia groundfish and sharks including XX species of rockfish, ... etc. (e.g., Figs 1 and 2). The full report, although not yet published through the Canadian Science Advisory Secretariat (CSAS), is available as a preprint at TODO. The visualizations for many stocks include mostly empty plots emphasizing species for which there remain many unknown unknowns. In other cases, nearly all the data available for data-rich stocks are visualized together on two pages for the first time.

<!-- The first page for each species features metadata and temporal and spatial trends and patterns from fisheries independent surveys and commercial catch and effort data. Specifically, it includes: -->

<!-- 1. Species-specific metadata (scientific and common names, taxonomy, links to FishBase and WoRMS species pages, current conservation status, most recent DFO stock assessment or other DFO document) -->
<!-- 1. Temporal and spatial population and fishery trends -->
<!-- 1. Relative biomass trends from fisheries independent surveys -->
<!-- 1. Maps of relative biomass from the most recent fisheries independent surveys -->
<!-- 1. Trends in fisheries catches through time by gear type and region -->
<!-- 1. Trends in bottom-trawl fisheries catch per unit effort (CPUE) through time by region -->
<!-- 1. Maps of fisheries CPUE by gear type for a recent window of time -->

<!-- The visualizations on the second page focus on biological samples of fish. Each page includes: -->

<!-- 1. Length frequencies by survey and year -->
<!-- 1. Age frequencies by survey and year -->
<!-- 1. Length age growth models -->
<!-- 1. von Bertalanffy growth curves fitted to age and length samples -->
<!-- 1. Weight-length growth models -->
<!-- 1. Age- and length-at-maturity models -->
<!-- 1. Maturity frequencies by maturity stage and month -->
<!-- 1. Counts of the number of available fish lengths, weights, maturities, ages, and aging structures from all surveys and all commercial samples in DFO databases -->

<!-- pcod, redbanded, rougheye? -->

# Discussion

We have described an automated report that presents up-to-date population trends and biological data from both commercial fisheries and DFO multi-species research surveys for groundfish in British Columbia, Canada. Stock assessments are not available for all species, or in all years for those that are assessed, but there are vast quantities of data collected in the fisheries and fishery-independent surveys. Our report is fully automated from data extraction to data processing to model fitting and finally to data visualization. TODO...

We designed the groundfish synopsis report to be fully automated from data extraction right through to report generation --- this was only possible because of the carefully organized and maintained relational databases housing DFO groundfish fishery and survey data. This report and its efficient data-to-document workflow demonstrate the value of investing in the development and ongoing maintenance of organized databases for any agency collecting and housing large quantities of data. Knowledge of and accessibility to available data within an agency is fundamentally important. The functions we created to extract groundfish data for the report are available within our organization for those requiring access to the databases, and facilitate internally reviewed, standardized, and consistent data extraction methods.

<!-- - give examples of what happens when databases are well organized and open (RAM Legacy, GPDD, ...) -->

Unlike detailed species-specific analyses conducted in stock assessments, we liken our report to window shopping. Our report gives a quick flavour of the available data for a given stock and hints at what trends for various indexes. However, the outputs in this report are not a substitute for species-specific stock assessment. For example, stock assessments may use different data sources than are presented in the report, and necessarily analyze the data on a stock-by-stock basis and make considerations for special cases which we do not. Biomass indices based on survey or commercial CPUE data in the report require careful, species-specific interpretation as they may not accurately represent abundance trends for some stocks. For example, biomass indices may be based on surveys with particularly high variability or low catchability for certain species and there are many reasons why commercial CPUE data may not be proportional to stock abundance [e.g., @harley2001].

<!-- Catch history in the report reflects recorded data and may not match catch time series reconstructions in stock assessments, which account for stock-specific considerations [e.g., @haigh2011]. Note also that fluctuations in catch may not reflect stock abundance and may be due to other factors including implementation of management measures. Discards shown in the catch plots only include bottom trawl discard weights from 1996 to present and trap, hook and line, midwater trawl and Strait of Georgia bottom trawl discards weights from 2006 to present due to low reliability of reported discards prior to implemetation of 100% observer coverage in the respective fisheries.  -->

Comparison to even tangentially related efforts

- mention data dashboards and their advantages and disadvantages
- reference Chris Rooper's Alaska work add Dan's work on the east coast

Where else this would be useful

- nearly anywhere with lots of populations or processes to monitor and available data maintained in an organized fashion
- especially useful for agencies responsible for a lot of stocks without the capacity to assess as frequently as management or ecological theory would suggest is ideal
- could also be useful for monitoring environmental data, not just biological data, and not just fisheries

Conclusions and where we are going with this next

- this was a useful exercise for us; already become an indispensable reference for many people and the tools have helped conduct other assessments rapidly
- plan to update on an annual or biennial basis and examine larger structural changes on a less frequent basis
- plan to include more climate related visualization such as centre of gravity and mean depth through time (ha! where to fit this...?)
- ...

# Acknowledgements

We are grateful for the ongoing collaboration between DFO, commercial fishers,
First Nations, and non-governmental organizations, which has made possible the
collection of the valuable data that underlies this report. We thank
the project's steering committee (Greg Workman, Robyn Forrest, Dana Haggarty,
Chris Grandin, and Rob Kronlund) for invaluable input into the report design
and feedback throughout its production.
We thank Chris Rooper and Daniel Ricard for their thorough and helpful reviews.
We thank Mary Thiess and Kieran Forge for chairing and rapporteuring the review meeting, respectively.
We thank Christie Whelan for her
support initiating this project. We thank Norm Olsen, Maria Surry, and Malcolm
Wyeth for providing support on data extraction and general database structure
and content.

\clearpage

# Figures

```{r page1, fig.cap = "Figure caption goes here."}
# include_graphics(here("figs/decades-and-booms-viridis.pdf"))
```

```{r page2, fig.cap = "Figure caption goes here."}
# include_graphics(here("figs/decades-and-booms-viridis.pdf"))
```

\clearpage

# References
